---
#title: "qm"

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#library
```{r}
#install.packages("acs")
#install.packages("tidycensus")
#install.packages("tidyverse")
#install.packages("dplyr")
#install.packages("readxl")
#install.packages("ggplot2")
#install.packages("scales")
#install.packages("readxl")
library(XML)
library(stringr)
library(acs)
library(tidycensus)
library(tidyverse)
library(dplyr)
library(readxl)
library(ggplot2)
library(scales)
library(readxl)
```
Sys.setenv(CENSUS_KEY = 'XX')
#Population data - VA state extracted from Census ACS 2015-2021 - Tract and County
```{r}
data <- data.frame()
years <- c(2017, 2018, 2019, 2020, 2021)
geo <- c("tract", "county")
for (i in years) {
  for (g in geo) {
  data_census <- get_acs(geography = g,
                       year = i,
                       survey = "acs5",
                       state = 'VA',
                       variables = c( adult_pop = 'S2301_C01_001E',
                                      unemployed='S2301_C04_001E',
                                      occupancy_all='B25014_001E',
                                      occupant_owner_1_to_1half='B25014_005E',
                                      occupant_owner_greater_1half='B25014_006E',
                                      occupant_owner_greater_2='B25014_007E',
                                      occupant_renter_1_to_1half='B25014_011E',
                                      occupant_renter_greater_1half='B25014_012E',                                      
                                      occupant_renter_greater_2='B25014_013E',  
                                      all_units = 'S2502_C01_001E',
                                      rent_units='S2502_C05_001E',
                                      households_total='B25044_001E',
                                      households_owner_no_vehicles='B25044_003E',
                                      households_renter_no_vehicles='B25044_010E'), 
                       geometry = FALSE,
                       output = 'wide' )
data_census$year = i
data_census$region_type = g
data <- rbind(data, data_census)

  }}


data <- data[, c("GEOID","year","adult_pop", "unemployed","occupancy_all", "occupant_owner_1_to_1half","occupant_owner_greater_1half","occupant_owner_greater_2","occupant_renter_1_to_1half", "occupant_renter_greater_1half","occupant_renter_greater_2", "all_units", "rent_units", "households_total", "households_owner_no_vehicles","households_renter_no_vehicles", "region_type")]

```


#Aggregating the results to Health Districts
```{r}
county <- data_census




health_district <- read_csv("~/git/sdc.geographies_dev/VA/State Geographies/Health Districts/2020/data/distribution/va_ct_to_hd_crosswalk.csv")
health_district <- merge(x = health_district, y = county, by.x = "ct_geoid", by.y = "GEOID")


health_district <- health_district %>%
  group_by(hd_geoid, year) %>%
  summarise(
    adult_pop = sum(adult_pop, na.rm = TRUE),
    unemployed = sum(unemployed, na.rm = TRUE),
    occupancy_all = sum(occupancy_all, na.rm = TRUE),
    occupant_owner_1_to_1half = sum(occupant_owner_1_to_1half, na.rm = TRUE),
    occupant_owner_greater_1half = sum(occupant_owner_greater_1half, na.rm = TRUE),
    occupant_owner_greater_2 = sum(occupant_owner_greater_2, na.rm = TRUE),
    occupant_renter_1_to_1half = sum(occupant_renter_1_to_1half, na.rm = TRUE),
    occupant_renter_greater_1half = sum(occupant_renter_greater_1half, na.rm = TRUE),
    occupant_renter_greater_2 = sum(occupant_renter_greater_2, na.rm = TRUE),
    all_units = sum(all_units, na.rm = TRUE),
    rent_units = sum(rent_units, na.rm = TRUE),
    households_total = sum(households_total, na.rm = TRUE),
    households_owner_no_vehicles = sum(households_owner_no_vehicles, na.rm = TRUE),
    households_renter_no_vehicles = sum(households_renter_no_vehicles, na.rm = TRUE),
  )

health_district <- rename(health_district, GEOID = hd_geoid)
health_district$region_type <- 'health district'
  
health_district <- health_district[, c("GEOID","year","adult_pop", "unemployed","occupancy_all", "occupant_owner_1_to_1half","occupant_owner_greater_1half","occupant_owner_greater_2","occupant_renter_1_to_1half", "occupant_renter_greater_1half","occupant_renter_greater_2", "all_units", "rent_units", "households_total", "households_owner_no_vehicles","households_renter_no_vehicles", "region_type")]

data <- rbind(data, health_district)

```







#Producing % unempolyment, non-car ownership, non-home ownership, and overcrowding
```{r}
# unemployment rate 
data$unemployed <- data$unemployed/data$adult_pop
data$unemployed  <- log(data$unemployed + 1)
data$unemployed <- as.numeric(data$unemployed)

#% households without a car 
data$noncar_ownership  <- (data$households_renter_no_vehicles+data$households_owner_no_vehicles) / data$households_total
data$noncar_ownership <- as.numeric(data$noncar_ownership)

#% households renting
data$nonhome_ownership  <- data$rent_units/data$all_units
data$nonhome_ownership <- as.numeric(data$nonhome_ownership)

# overcrowded households 
data$household_overcrowding <- (
  data$occupant_owner_1_to_1half +  
  data$occupant_owner_greater_1half + 
  data$occupant_owner_greater_2 + 
  data$occupant_renter_1_to_1half +  
  data$occupant_renter_greater_1half + 
  data$occupant_renter_greater_2
) / (data$occupancy_all)

data$household_overcrowding  <- log(1 + (data$household_overcrowding)) 
data$household_overcrowding <- as.numeric(data$household_overcrowding)


townsend_data <- subset(data, select = c(GEOID, region_type, unemployed, noncar_ownership, nonhome_ownership, household_overcrowding, year))
```



#Scaling the variables for each year and region type 
```{r}

townsend_data <- townsend_data %>%
  group_by(region_type, year) %>%
  mutate(across(c(unemployed, noncar_ownership,nonhome_ownership,household_overcrowding ), ~scale(.), .names="z_{.col}")) %>% ungroup()

```


#Producing the Townsend Dep Index and requested file 
```{r}
townsend_data$Townsend_Index_z_sum <- townsend_data$z_unemployed + townsend_data$z_noncar_ownership + townsend_data$z_nonhome_ownership + townsend_data$z_household_overcrowding
townsend_data <- townsend_data %>%   group_by(region_type, year) %>%
  mutate(Townsend_Index_z_sum_z = scale(Townsend_Index_z_sum)) %>% ungroup()

townsend_data$Townsend_Index_z_sum_z <- (townsend_data$Townsend_Index_z_sum_z) * -1

townsend_data <- townsend_data %>%
  group_by(region_type, year) %>%
  mutate(
    Townsend_Index_z_standardized = (Townsend_Index_z_sum - min(Townsend_Index_z_sum)) /
                                     (max(Townsend_Index_z_sum) - min(Townsend_Index_z_sum))
  ) %>%
  ungroup()



townsend_data$measure = 'material_deprivation_indicator'
townsend_data <- rename(townsend_data, value = Townsend_Index_z_standardized, geoid = GEOID)



townsend_data <- townsend_data %>%
  group_by(year, region_type) %>%
  mutate(value = rescale(value)) %>%
  ungroup()





townsend_bi <- subset(townsend_data, select = c(geoid, measure, year, region_type, value))
```

townsend_bi <- townsend_bi %>% filter(year==2020)

merged_df <- merge(townsend_bi, HOI_V3_14_Variables_For_UVA, by.x = "geoid", by.y = "FIPS")



ggplot(merged_df, aes(x=value, y=Townsend)) + 
  geom_point(alpha=0.7, color="blue") +
  geom_abline(intercept=0, slope=1, color="red", linetype="solid", size=1) + 
  ggtitle("UVA vs VDH - Townsend") +
  xlab("UVA") +
  ylab("VDH")

print(colnames(merged_df))

#Write CSV
```{r}

readr::write_csv(townsend_bi_all, xzfile("../../data/distribution/va_hdcttr_vdh_2015_2021_material_deprivation_index.csv.xz", compression = 9))

```

