---
title: "catchment_scores - urgent care - VA"
author: "--"
date: "2/4/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Libraries
```{r}
library(sf)
library(tidygeocoder)
library(dplyr)
library(community)
library(readr)
```
# Start new
```{r}
#start clean
rm(list=ls())
```
# Working directory
```{r}
setwd("~/VDH/Floating_Catchment_Areas/va/5. hospitals_va_model")
```

# List of buffer counties for VA
```{r}
source("~/VDH/Floating_Catchment_Areas/va/buffer_counties_va.R")
```
# Read data
```{r}
#source: HIFLD
hosp_usa <- read.csv("~/test_symlink/hospitals/Hospitals_usa.csv")
hosp_va <- hosp_usa %>% filter(STATUS=='OPEN', STATE=='VA')
#hosp_va_acute <- hosp_va %>% filter(TYPE == 'GENERAL ACUTE CARE') 
hosp_va <- hosp_va %>% filter(TYPE != 'MILITARY' , TYPE != 'PSYCHIATRIC' , TYPE != 'REHABILITATION') 

hosp_va_buffer <- hosp_usa %>% filter(STATUS=='OPEN', STATE=='VA' | COUNTYFIPS %in% buffer_counties_state$identifier ) 

#acute care
provider <- hosp_va_buffer %>% filter(TYPE == 'GENERAL ACUTE CARE' | TYPE == 'CRITICAL ACCESS') %>% filter(BEDS > 0)


#change names lat lon
names(provider)[names(provider)=='LATITUDE'] <- 'lat'
names(provider)[names(provider)=='LONGITUDE'] <- 'lon'



#Save list of hospitals:93 in VA
write_csv(provider, "general_acute_care/va_hifld_2021_hospitals.csv")
#provider <- read.csv("hospitals_ncr.csv", row.names = 1)
```

# Fix data: cluster providers by lat-lon to have unique geographical providers (e.g. 15 doctors in one location)
```{r}
## collapse by location
provider$doctors <- 1
provider$location <- paste0(provider$lat, ",", provider$lon)
counts <- tapply(provider$doctors, provider$location, sum)
locations <- which(!duplicated(provider$location))
provider <- provider[locations,]
provider$doctors <- counts[provider$location]

## assign IDs just to be explicit
provider$ID <- paste0("l", seq_len(nrow(provider)))
```

# Data combined: population
```{r}
library(tidycensus)
library(tidyverse)

census_api_key("eba406410c653b81d6a795ac4e989221f7bdf302")

# Bring in census tract data. 
pop_va_hosp <- get_acs(geography = "block group", 
                        year = 2020,
                        variables = c(population = "B01003_001E",
                                      medinc = "B19013_001E"
                        ),
                        state = "VA",
                        survey = "acs5",
                        output = "wide",
                        geometry = TRUE)

pop_dc_hosp <- get_acs(geography = "block group", 
                        year = 2020,
                        variables = c(population = "B01003_001E",
                                      medinc = "B19013_001E"
                        ),
                        state = "DC",
                        survey = "acs5",
                        output = "wide",
                        geometry = TRUE)

pop_md_hosp <- get_acs(geography = "block group", 
                        year = 2020,
                        variables = c(population = "B01003_001E",
                                      medinc = "B19013_001E"
                        ),
                        state = "MD",
                         county = buffer_counties_state$NAMELSAD[buffer_counties_state$STUSPS=='MD'],
                        survey = "acs5",
                        output = "wide",
                        geometry = TRUE)

pop_pa_hosp <- get_acs(geography = "block group", 
                        year = 2020,
                        variables = c(population = "B01003_001E",
                                      medinc = "B19013_001E"
                        ),
                        state = "PA",
                         county = buffer_counties_state$NAMELSAD[buffer_counties_state$STUSPS=='PA'],
                        survey = "acs5",
                        output = "wide",
                        geometry = TRUE)

pop_wv_hosp <- get_acs(geography = "block group", 
                        year = 2020,
                        variables = c(population = "B01003_001E",
                                      medinc = "B19013_001E"
                        ),
                        state = "WV",
                         county = buffer_counties_state$NAMELSAD[buffer_counties_state$STUSPS=='WV'],
                        survey = "acs5",
                        output = "wide",
                        geometry = TRUE)

pop_ky_hosp <- get_acs(geography = "block group", 
                        year = 2020,
                        variables = c(population = "B01003_001E",
                                      medinc = "B19013_001E"
                        ),
                        state = "KY",
                         county = buffer_counties_state$NAMELSAD[buffer_counties_state$STUSPS=='KY'],
                        survey = "acs5",
                        output = "wide",
                        geometry = TRUE)

pop_tn_hosp <- get_acs(geography = "block group", 
                        year = 2020,
                        variables = c(population = "B01003_001E",
                                      medinc = "B19013_001E"
                        ),
                        state = "TN",
                         county = buffer_counties_state$NAMELSAD[buffer_counties_state$STUSPS=='TN'],
                        survey = "acs5",
                        output = "wide",
                        geometry = TRUE)

pop_nc_hosp <- get_acs(geography = "block group", 
                        year = 2020,
                        variables = c(population = "B01003_001E",
                                      medinc = "B19013_001E"
                        ),
                        state = "NC",
                         county = buffer_counties_state$NAMELSAD[buffer_counties_state$STUSPS=='NC'],
                        survey = "acs5",
                        output = "wide",
                        geometry = TRUE)

pop_de_hosp <- get_acs(geography = "block group", 
                        year = 2020,
                        variables = c(population = "B01003_001E",
                                      medinc = "B19013_001E"
                        ),
                        state = "DE",
                         county = buffer_counties_state$NAMELSAD[buffer_counties_state$STUSPS=='DE'],
                        survey = "acs5",
                        output = "wide",
                        geometry = TRUE)

#joint df
pop_buffer_hosp <- rbind(pop_va_hosp, pop_dc_hosp, pop_de_hosp, pop_ky_hosp, pop_md_hosp, pop_nc_hosp, pop_pa_hosp, pop_tn_hosp, pop_wv_hosp)

#
va <- pop_buffer_hosp %>% select(GEOID, NAME, population=population )


# data combined
#centroid and coordinates

data_combined <- data.frame(
  GEOID = va$GEOID,
  population = va$population,
sf::st_coordinates(st_centroid(va$geometry))
)

data_combined <- data_combined %>% filter(!is.na(data_combined$X))

data_combined <- data_combined %>% filter(GEOID != 517000323001 )
data_combined <- data_combined %>% filter(GEOID != 517000323002 )
data_combined <- data_combined %>% filter(GEOID != 517000323003 )

```


# Travel time
```{r}
library(osrm)
options(osrm.server = "http://104.248.112.16:5001/", osrm.profile = "car")

if(!file.exists("traveltimes_exercise.csv")){
  
  #provider <- st_drop_geometry( provider )
    
  traveltimes <- osrmTable(
    src = data_combined[, c( "X", "Y")],  #population-demand
    dst =  provider[, c( "lon", "lat")]  #providers supply
  )$duration
  
  names(traveltimes) <-   provider[, c("ID")] 
  
  write.csv(
    cbind(GEOID = data_combined[, c("GEOID")], as.data.frame(traveltimes)),
    "traveltimes_exercise.csv", row.names = FALSE
  )
}

traveltimes <- read.csv("traveltimes_exercise.csv", row.names = 1)
names(traveltimes) <-   provider[, c("ID")] 

```


# Define geography id - Geography-GEOID from initial file may be outdated
```{r}
library(tigris)
library(maps)
library(sf)
# add block geoids
# get US blocks shapefile
blocks_VA <- st_as_sf(block_groups(state="VA", year=2020), crs = 4326) #, year=2010
blocks_DC <- st_as_sf(block_groups(state= "DC", year=2020), crs = 4326) #, year=2010
blocks_MD <- st_as_sf(block_groups(state= "MD", year=2020), crs = 4326) #, year=2010
blocks_NC <- st_as_sf(block_groups(state= "NC", year=2020), crs = 4326) #, year=2010
blocks_DE <- st_as_sf(block_groups(state= "DE", year=2020), crs = 4326) #, year=2010
blocks_TN <- st_as_sf(block_groups(state= "TN", year=2020), crs = 4326) #, year=2010
blocks_WV <- st_as_sf(block_groups(state= "WV", year=2020), crs = 4326) #, year=2010
blocks_KY <- st_as_sf(block_groups(state= "KY", year=2020), crs = 4326) #, year=2010
blocks_PA <- st_as_sf(block_groups(state= "PA", year=2020), crs = 4326) #, year=2010

blocks <- rbind( blocks_VA, blocks_DC, blocks_MD , blocks_NC , blocks_DE , blocks_TN , blocks_WV , blocks_KY , blocks_PA) 
blocks <- st_transform(blocks, 4326)



# blocks_DC <- readRDS("~/VDH/Floating_Catchment_Areas/dmv/blocks_DC_2014" )
# blocks_MD <- readRDS( "~/VDH/Floating_Catchment_Areas/dmv/blocks_MD_2014" )
#blocks_VA <- readRDS("~/VDH/Floating_Catchment_Areas/dmv/blocks_VA_2014" )

# blocks <- rbind(blocks_VA, blocks_MD)
# blocks <- rbind(blocks, blocks_DC)

#blocks <- blocks_VA
# lon and lat to geo-points

geopts <- provider %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326)

geopts <- st_transform(geopts, 4326)

# indeces of bgs which contain a geopoint
inds <- st_within(geopts$geometry, blocks$geometry, sparse=T)
blk_list <- c()
for (i in inds){
  if (identical(blocks$NAME[i],character(0))){
    blk_list<- append(blk_list, NA)}
  else{
    blk_list <- append(blk_list, blocks$GEOID[i])}
}
provider['geoid_blk_new'] <- blk_list


# This assumes that you have longitude and latitude of each location and if you want bg or tract or county just replace 'blocks_VA <- st_as_sf(blocks(state="VA"))' with a corresponding value, e.g. 'tracts_VA <- st_as_sf(tracts(state="VA"))'. I believe it uses either tigris or maps package

```

# Add count of providers per geography using matching codes
```{r}

#providers w geoid
provider$GEOID <- substr(provider$geoid_blk_new, 1, 12 ) 
#data_combined with geoid: ok
num_providers <- provider %>% group_by(GEOID) %>% summarise(prov_cnt = sum(doctors) )

#join providers to block groups
data_combined$GEOID <-  as.character( data_combined$GEOID)
data_combined <- data_combined %>% left_join(num_providers, by= "GEOID" )

```
# Add mean and median of 10 nearest drive times 
```{r}
#mean of 10 nearest
top_mean <- function(x) {  
   mean(head(sort(x ), 10) ) }
#median of 10 nearest
top_median <- function(x) {  
   median(head(sort(x ), 10) ) }
#apply rowwise
traveltimes_near <- data.frame(near_10_mean=apply(traveltimes, 1, top_mean), 
                               near_10_median=apply(traveltimes, 1, top_median)) 
#rownames_to_column(traveltimes_near, var = "GEOID")
traveltimes_near$GEOID <- row.names(traveltimes_near) 
#join mean median traveltimes to geographies
data_combined <- data_combined %>% left_join(traveltimes_near, by= "GEOID")
```

# Prepara data to save

```{r}
#raw traveltimes: traveltimes matrix already estimated and with colnames arranged 
traveltimes <- read.csv("traveltimes_exercise.csv", row.names = 1)
names(traveltimes) <-   provider[, c("ID")] 
#population: always recheck relevant population: ie. for pediatrics: pop 0-17 years
population <- data_combined %>% select(GEOID, population, prov_cnt, near_10_mean, near_10_median, X, Y)
# realign travel times
traveltimes <- traveltimes[as.character(population$GEOID), provider$ID]

```

# Save files to estimate catchment areas
```{r}
write.csv(provider[, c("ID", address="ADDRESS", "lat", "lon", "doctors", beds="BEDS")] , "general_acute_care/provider_20.csv", row.names = FALSE)
write.csv(cbind(GEOID = rownames(traveltimes), traveltimes), "general_acute_care/traveltimes_trimmed_20.csv", row.names = FALSE)
write.csv(population, "general_acute_care/population.csv", row.names = FALSE)
```



