---
title: "catchment_scores"
author: "--"
date: "2/4/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# libraries
```{r}
library(catchment)
library(sf)
library(leaflet)
library(dplyr)
```
# start new
```{r}
 rm(list=ls())
```
#working directory
```{r}
setwd("~/VDH/Floating_Catchment_Areas/dmv/5.1 hospitals_dmv_model_2015-2022/5.1 hospitals_dmv_model_2016")
getwd()
```
# read data
```{r}
year_spec <- 2016

population <- read.csv("population.csv")
provider <- read.csv("provider.csv")
rownames(provider) <- provider$ID
traveltimes <- read.csv("traveltimes_trimmed.csv", row.names = 1)
```
# prepare map
```{r}
#blockgroup_shapes <- st_transform(read_sf("blockgroups.geojson", as_tibble = FALSE), 4326)
blockgroup_shapes_va <- st_transform( tigris::block_groups(state = c("VA"), class = "sf", year=year_spec), 4326)
blockgroup_shapes_dc <- st_transform( tigris::block_groups(state = c("DC"), class = "sf", year=year_spec), 4326)
blockgroup_shapes_md <- st_transform( tigris::block_groups(state = c("MD"), class = "sf", year=year_spec), 4326)
blockgroup_shapes_pa <- st_transform( tigris::block_groups(state = c("PA"), class = "sf", year=year_spec), 4326)
blockgroup_shapes_wv <- st_transform( tigris::block_groups(state = c("WV"), class = "sf", year=year_spec), 4326)
 
blockgroup_shapes_tot <- rbind(blockgroup_shapes_va, blockgroup_shapes_dc ,blockgroup_shapes_md, blockgroup_shapes_pa, blockgroup_shapes_wv )
geoid_sample <- population$GEOID
blockgroup_shapes <- blockgroup_shapes_tot %>% filter(GEOID %in% geoid_sample)

rownames(blockgroup_shapes) <- blockgroup_shapes$GEOID
blockgroup_shapes <- blockgroup_shapes[as.character(population$GEOID),]
#
# rownames(blockgroup_shapes_va) <- blockgroup_shapes_va$GEOID
# blockgroup_shapes_va <- blockgroup_shapes_va[as.character(population$GEOID),]
# #dc
# rownames(blockgroup_shapes_dc) <- blockgroup_shapes_dc$GEOID
# blockgroup_shapes_dc <- blockgroup_shapes_dc[as.character(population$GEOID),]

##new
shape_counties <- tigris::counties(state = c("VA", "MD", "DC"), class = "sf", year=year_spec)
shape_counties_dmv <- shape_counties %>% filter(GEOID == 51013 | GEOID ==51059 | GEOID == 51600 | GEOID == 51107 | GEOID == 51610 | GEOID == 51683 | 
                                                  GEOID == 51685 | GEOID == 51153 | GEOID == 51510 | GEOID == 11001 | GEOID ==  24021 | GEOID == 24031 | GEOID == 24033 | GEOID == 24017) 
#format for counties
shape_counties_dmv <- st_transform(shape_counties_dmv, 4326 )

#blockgroup_shapes <- sf::st_as_sf(blockgroup_shapes)
#create map
map <- leaflet(blockgroup_shapes, options = leafletOptions(attributionControl = FALSE)) %>%
  addTiles() %>%
  addScaleBar("bottomleft") %>%
  addMapPane("lines", zIndex = 410) %>%
  addMapPane("points", zIndex = 411) %>%
  addPolygons(
    fillColor = colorNumeric("RdYlBu", population$population)(population$population),
    fillOpacity = 1, stroke = FALSE, group = "Population", label = population$population
  ) %>%
  hideGroup("Population") %>%
  addLayersControl(
    position = "topleft", overlayGroups = c("Pediatrics", "Population", "Access")
  ) %>%
  addCircles(
    data = provider, color = "blue", lng = ~lon, lat = ~lat, weight = 4,
    label = ~ paste0("ID: ", ID)
  )




```
# calculate and display catchment areas
#Revise titles of figures
## 2-step
```{r}
population$providers_2sfca <- catchment_ratio(
  population, provider, traveltimes, 30,
  consumers_value = "population", providers_id = "ID", providers_value = "doctors", verbose = TRUE
) * 100000

pal <- colorBin("RdYlBu", population$providers_2sfca, bins = 6)

plot_2sfca <- map %>%
  addControl(paste("Hospitals Per 100,000 People (2-Step Floating Catchment Area) in", year_spec), "topright") %>%
  addLegend("bottomright", pal, population$providers_2sfca, opacity = .7) %>%
  addPolygons(
    fillColor = pal(population$providers_2sfca), fillOpacity = .7, weight = 1, color = "#000",
    highlightOptions = highlightOptions(color = "#fff"), group = "Access",
    label = paste0(
      "GEOID: ", population$GEOID, ", Population Ages 0-17: ", population$population,
      ", Per 1k People: ", round(population$providers_2sfca, 4), ", In Region: ",
      round(population$population * population$providers_2sfca / 1000, 4)
    )
  ) %>%
  addPolylines(data = shape_counties_dmv, color = "black", opacity = 1, weight = 2.5) %>%
  addCircles(
    data = provider, color = "blue", lng = ~lon, lat = ~lat, weight = 4,
    label = ~ paste0("ID: ", ID)
  )

## enhanced 2-step

weight <- list(c(60, .042), c(30, .377), c(20, .704), c(10, .962))
population$providers_e2sfca <- catchment_ratio(
  population, provider, traveltimes, weight,
  consumers_value = "population", providers_id = "ID", providers_value = "doctors", verbose = TRUE
) * 100000

pal <- colorBin("RdYlBu", population$providers_e2sfca, bins = 6)

plot_e2sfca <- map %>%
  addControl(paste("Hospitals Per 100,000 People (enhanced 2-Step Floating Catchment Area) in", year_spec), "topright") %>%
  addLegend("bottomright", pal, population$providers_e2sfca, opacity = .7) %>%
  addPolygons(
    fillColor = pal(population$providers_e2sfca), fillOpacity = .7, weight = 1, color = "#000",
    highlightOptions = highlightOptions(color = "#fff"), group = "Access",
    label = paste0(
      "GEOID: ", population$GEOID, ", Population Ages 0-17: ", population$population,
      ", Per 1k People: ", round(population$providers_e2sfca, 4), ", In Region: ",
      round(population$population * population$providers_e2sfca / 1000, 4)
    )
  ) %>% addPolylines(data = shape_counties_dmv, color = "black", opacity = 1, weight = 2.5) %>%
  addCircles(
    data = provider, color = "blue", lng = ~lon, lat = ~lat, weight = 4,
    label = ~ paste0("ID: ", ID)
  )


## 3-step

population$providers_3sfca <- catchment_ratio(
  population, provider, traveltimes, "gaussian", normalize_weight = TRUE, scale = 20,
  consumers_value = "population", providers_id = "ID", providers_value = "doctors", verbose = TRUE
) * 100000

pal <- colorBin("RdYlBu", population$providers_3sfca, bins = 6)

plot_3sfca <- map %>%
  addControl( paste("Hospitals Per 100,000 People (3-Step Floating Catchment Area) in", year_spec) , "topright") %>%
  addLegend("bottomright", pal, population$providers_3sfca, opacity = .7) %>%
  addPolygons(
    fillColor = pal(population$providers_3sfca), fillOpacity = .7, weight = 1, color = "#000",
    highlightOptions = highlightOptions(color = "#fff"), group = "Access",
    label = paste0(
      "GEOID: ", population$GEOID, ", Population Ages 0-17: ", population$population,
      ", Per 1k People: ", round(population$providers_3sfca, 4), ", In Region: ",
      round(population$population * population$providers_3sfca / 1000, 4)
    )
  ) %>% addPolylines(data = shape_counties_dmv, color = "black", opacity = 1, weight = 2.5) %>%
  addCircles(
    data = provider, color = "blue", lng = ~lon, lat = ~lat, weight = 4,
    label = ~ paste0("ID: ", ID)
  )
```
#plot summary
```{r}
# plot_2sfca
# plot_e2sfca
# plot_3sfca
```
# Save pgadmin
```{r}
######
#reshape wide-long
#DEFINE relevant population and names
dmv_bg_2021_acccess_scores_wide <- population   %>% select( geoid="GEOID", 
                                                            hosp_pop_cnt ="population",
                                                            hosp_cnt = "hosp_cnt",
                                                            #hosp_beds_cnt = "beds_cnt",
                                                            hosp_near_10_mean = "near_10_mean",
                                                            hosp_near_10_median = "near_10_median",
                                                            hosp_2sfca = "providers_2sfca" ,
                                                            hosp_e2sfca = "providers_e2sfca", 
                                                            hosp_3sfca = "providers_3sfca" )
```

# long format block group
```{r}
dmv_bg_2021_acccess_scores_bg <- tidyr::gather(dmv_bg_2021_acccess_scores_wide, measure, value, 
                                            hosp_pop_cnt,
                                            hosp_cnt,
                                            #hosp_beds_cnt, 
                                            hosp_near_10_mean,
                                            hosp_near_10_median,
                                            hosp_2sfca, 
                                            hosp_e2sfca, 
                                            hosp_3sfca)
#define year
dmv_bg_2021_acccess_scores_bg$year <- year_spec
#define measure type
dmv_bg_2021_acccess_scores_bg$measure_type <- as.factor(ifelse(dmv_bg_2021_acccess_scores_bg$measure=="hosp_2sfca" |
                                                            dmv_bg_2021_acccess_scores_bg$measure=="hosp_e2sfca" |
                                                              dmv_bg_2021_acccess_scores_bg$measure=="hosp_3sfca", 'index','count') )
```

#include name: geographic name for the smallest geographic unit: block group in this case
```{r}
#a) connect
con <- RPostgreSQL::dbConnect(drv = RPostgreSQL::PostgreSQL(),
                              dbname = "sdad",
                              host = "postgis1",
                              port = 5432,
                              user = Sys.getenv(x = "DB_USR"),
                              password = Sys.getenv(x = "DB_PWD"))

#b) query
name_geo <- sf::st_read(con, 
  query= "
SELECT  *
FROM dc_geographies.ncr_cttrbg_tiger_2010_2020_geo_names "
)

#join region_name + region_type
dmv_bg_2021_acccess_scores_bg$geoid <- as.character(dmv_bg_2021_acccess_scores_bg$geoid)
dmv_bg_2021_acccess_scores_bg <- dmv_bg_2021_acccess_scores_bg %>% left_join(name_geo, by="geoid")
#order
dmv_bg_2021_acccess_scores_bg <- dmv_bg_2021_acccess_scores_bg %>% select("geoid","region_type","region_name", "measure", "value", "year", "measure_type")
######
```

# AGGREGATION tract
```{r}
#generate id tract 
#the relevant file is WIDE
dmv_bg_2021_acccess_scores_wide_aggreg <- dmv_bg_2021_acccess_scores_wide  %>% mutate(tract_geoid = substr(geoid,1,11))
#add region name - tract
# dmv_bg_2021_acccess_scores_wide_aggreg <- dmv_bg_2021_acccess_scores_wide_aggreg %>% left_join(name_geo, by= c("tract_geoid"= "geoid") )

## measures tract
dmv_bg_2021_acccess_scores_tract_wide <- dmv_bg_2021_acccess_scores_wide_aggreg %>% group_by(tract_geoid) %>% summarise(
  hosp_pop = sum(hosp_pop_cnt, na.rm = TRUE), 
  hosp_cnt = sum(hosp_cnt, na.rm = TRUE), 
  #hosp_beds_cnt =  sum(hosp_beds_cnt, na.rm = TRUE), 
  hosp_near_10_mean = weighted.mean(hosp_near_10_mean, hosp_pop_cnt, na.rm = TRUE), 
  hosp_near_10_median = weighted.mean(hosp_near_10_median, hosp_pop_cnt, na.rm = TRUE), 
  hosp_2sfca = weighted.mean(hosp_2sfca, hosp_pop_cnt, na.rm = TRUE), 
  hosp_e2sfca = weighted.mean(hosp_e2sfca, hosp_pop_cnt, na.rm = TRUE),
  hosp_3sfca = weighted.mean(hosp_3sfca, hosp_pop_cnt, na.rm = TRUE),
  )

               
```

# long format tract
```{r}
dmv_bg_2021_acccess_scores_tract <- tidyr::gather(dmv_bg_2021_acccess_scores_tract_wide, measure, value, 
                                            hosp_pop,
                                            hosp_cnt,
                                            #hosp_beds_cnt, 
                                            hosp_near_10_mean,
                                            hosp_near_10_median,
                                            hosp_2sfca, 
                                            hosp_e2sfca, 
                                            hosp_3sfca)
#define year
dmv_bg_2021_acccess_scores_tract$year <- year_spec
#define measure type
dmv_bg_2021_acccess_scores_tract$measure_type <- as.factor(ifelse(dmv_bg_2021_acccess_scores_tract$measure=="hosp_2sfca" |
                                                            dmv_bg_2021_acccess_scores_tract$measure=="hosp_e2sfca" |
                                                              dmv_bg_2021_acccess_scores_tract$measure=="hosp_3sfca", 'index','count') )
```

#include name: geographic name for geographic unit: tract in this case
```{r}
#join region_name + region_type
dmv_bg_2021_acccess_scores_tract$tract_geoid <- as.character(dmv_bg_2021_acccess_scores_tract$tract_geoid)
dmv_bg_2021_acccess_scores_tract <- dmv_bg_2021_acccess_scores_tract %>% left_join(name_geo, by= c("tract_geoid" = "geoid") )
#order
dmv_bg_2021_acccess_scores_tract <- dmv_bg_2021_acccess_scores_tract %>% select(geoid="tract_geoid","region_type","region_name", "measure", "value", "year", "measure_type")
```


# AGGREGATION county
```{r}
#generate id tract 
#the relevant file is WIDE
dmv_bg_2021_acccess_scores_wide_aggreg <- dmv_bg_2021_acccess_scores_wide  %>% mutate(county_geoid = substr(geoid,1,5))
#add region name - tract
# dmv_bg_2021_acccess_scores_wide_aggreg <- dmv_bg_2021_acccess_scores_wide_aggreg %>% left_join(name_geo, by= c("tract_geoid"= "geoid") )

## measures tract
dmv_bg_2021_acccess_scores_county_wide <- dmv_bg_2021_acccess_scores_wide_aggreg %>% group_by(county_geoid) %>% summarise(
  hosp_pop = sum(hosp_pop_cnt, na.rm = TRUE), 
  hosp_cnt = sum(hosp_cnt, na.rm = TRUE), 
  #hosp_beds_cnt =  sum(hosp_beds_cnt, na.rm = TRUE), 
  hosp_near_10_mean = weighted.mean(hosp_near_10_mean, hosp_pop_cnt, na.rm = TRUE), 
  hosp_near_10_median = weighted.mean(hosp_near_10_median, hosp_pop_cnt, na.rm = TRUE), 
  hosp_2sfca = weighted.mean(hosp_2sfca, hosp_pop_cnt, na.rm = TRUE), 
  hosp_e2sfca = weighted.mean(hosp_e2sfca, hosp_pop_cnt, na.rm = TRUE),
  hosp_3sfca = weighted.mean(hosp_3sfca, hosp_pop_cnt, na.rm = TRUE),
  )

               
```

# long format tract
```{r}
dmv_bg_2021_acccess_scores_county <- tidyr::gather(dmv_bg_2021_acccess_scores_county_wide, measure, value, 
                                            hosp_pop,
                                            hosp_cnt,
                                            #hosp_beds_cnt, 
                                            hosp_near_10_mean,
                                            hosp_near_10_median,
                                            hosp_2sfca, 
                                            hosp_e2sfca, 
                                            hosp_3sfca)
#define year
dmv_bg_2021_acccess_scores_county$year <- year_spec
#define measure type
dmv_bg_2021_acccess_scores_county$measure_type <- as.factor(ifelse(dmv_bg_2021_acccess_scores_county$measure=="hosp_2sfca" |
                                                            dmv_bg_2021_acccess_scores_county$measure=="hosp_e2sfca" |
                                                              dmv_bg_2021_acccess_scores_county$measure=="hosp_3sfca", 'index','count') )
```

#include name: geographic name for geographic unit: county in this case
```{r}
#join region_name + region_type
dmv_bg_2021_acccess_scores_county$tract_geoid <- as.character(dmv_bg_2021_acccess_scores_county$county_geoid)
dmv_bg_2021_acccess_scores_county <- dmv_bg_2021_acccess_scores_county %>% left_join(name_geo, by= c("county_geoid" = "geoid") )
#order
dmv_bg_2021_acccess_scores_county <- dmv_bg_2021_acccess_scores_county %>% select(geoid="county_geoid","region_type","region_name", "measure", "value", "year", "measure_type")
```

# stack all the measures for BLOCK GROUP + TRACT + COUNTY: include NAME
```{r}
dmv_bg_2021_acccess_scores_hospitals <- rbind(dmv_bg_2021_acccess_scores_bg, 
                                    dmv_bg_2021_acccess_scores_tract,
                                    dmv_bg_2021_acccess_scores_county)
```

#change one name
```{r}

dmv_bg_2021_acccess_scores_hospitals$measure <- ifelse(dmv_bg_2021_acccess_scores_hospitals$measure=='hosp_pop', 'hosp_pop_cnt' , dmv_bg_2021_acccess_scores_hospitals$measure)
```


```{r}
#SAVE with specific name
#The naming convention for data tables within a schema is as follows: 
#<location>_<resolution>_<data source>_<time period>_<title> 

#dmv_bg_2021_acccess_scores_hospitals$year <- year_spec

write.csv(dmv_bg_2021_acccess_scores_hospitals, "ncr_bgtrct_2016_acccess_scores_hospitals.csv", row.names = FALSE )

```


