---
title: "catchment_scores - obgyn - VA"
author: "--"
date: "2/4/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#libraries
```{r}
library(sf)
library(tidygeocoder)
library(dplyr)
library(community)
library(readr)
library(tidyverse)
```
#clean df
```{r}
#start clean
rm(list=ls())
```
#working directory
```{r}
setwd("~/VDH/Floating_Catchment_Areas/va/1. obgyn_va_model/")
```
# load data provider
```{r}
#from pgadmin
#a) conn
con <- RPostgreSQL::dbConnect(drv = RPostgreSQL::PostgreSQL(),
                               dbname = "sdad",
                               host = "postgis1",
                               port = 5432,
                               user = Sys.getenv(x = "DB_USR"),
                               password = Sys.getenv(x = "DB_PWD"))

#b) query
provider  <- sf::st_read(
  con, 
  query= "
    SELECT *
  FROM dc_health_behavior_diet.va_pl_webmd_2021_doctors_obgyn"
)

#c) Disconnect
RPostgreSQL::dbDisconnect(con)

names(provider)[names(provider)=='latitude'] <- 'lat'
names(provider)[names(provider)=='longitude'] <- 'lon'
provider <- st_drop_geometry(provider)


```
# Fix data: cluster providers by lat-lon to have unique geographical providers (e.g. 15 doctors in one location)
```{r}
## collapse by location
provider$doctors <- 1
provider$location <- paste0(provider$lat, ",", provider$lon)
#identify unique values
provider <- provider %>% distinct(name, location, .keep_all = TRUE)

counts <- tapply(provider$doctors, provider$location, sum)
locations <- which(!duplicated(provider$location))
provider <- provider[locations,]
provider$doctors <- counts[provider$location]

## assign IDs just to be explicit
provider$ID <- paste0("l", seq_len(nrow(provider)))


provider <- provider %>% filter(!is.na(lat))

```

# List of buffer counties for VA
```{r}
source("~/VDH/Floating_Catchment_Areas/va/buffer_counties_va.R")
```

# Data combined: population
```{r}
library(tidycensus)
library(tidyverse)

census_api_key("eba406410c653b81d6a795ac4e989221f7bdf302")

pop_va_fem <- get_acs(geography = "block group", 
                  year = 2020,
                  variables = c(population = "B01001_001",
                                male= "B01001_002",
                                female = "B01001_026", 
                                FemaleUnder5years="B01001_027",
                                Female5to9years="B01001_028",
                                Female10to14years="B01001_029",
                                Female15to17years="B01001_030",
                                Female18and19years="B01001_031",
                                Female20years="B01001_032",
                                Female21years="B01001_033",
                                Female22to24years="B01001_034",
                                Female25to29years="B01001_035",
                                Female30to34years="B01001_036",
                                Female35to39years="B01001_037",
                                Female40to44years="B01001_038",
                                Female45to49years="B01001_039",
                                Female50to54years="B01001_040",
                                Female55to59years="B01001_041",
                                Female60and61years="B01001_042",
                                Female62to64years="B01001_043",
                                Female65and66years="B01001_044",
                                Female67to69years="B01001_045",
                                Female70to74years="B01001_046",
                                Female75to79years="B01001_047",
                                Female80to84years="B01001_048",
                                Female85yearsandover="B01001_049"),
                  state = c("VA", "DC"),
                  #state = c("VA", "DC", "MD", "NC", "TN", "WV", "KY"),
                  survey = "acs5",
                  output = "wide",
                  geometry = TRUE)

pop_md_fem <- get_acs(geography = "block group", 
                  year = 2020,
                  variables = c(population = "B01001_001",
                                male= "B01001_002",
                                female = "B01001_026", 
                                FemaleUnder5years="B01001_027",
                                Female5to9years="B01001_028",
                                Female10to14years="B01001_029",
                                Female15to17years="B01001_030",
                                Female18and19years="B01001_031",
                                Female20years="B01001_032",
                                Female21years="B01001_033",
                                Female22to24years="B01001_034",
                                Female25to29years="B01001_035",
                                Female30to34years="B01001_036",
                                Female35to39years="B01001_037",
                                Female40to44years="B01001_038",
                                Female45to49years="B01001_039",
                                Female50to54years="B01001_040",
                                Female55to59years="B01001_041",
                                Female60and61years="B01001_042",
                                Female62to64years="B01001_043",
                                Female65and66years="B01001_044",
                                Female67to69years="B01001_045",
                                Female70to74years="B01001_046",
                                Female75to79years="B01001_047",
                                Female80to84years="B01001_048",
                                Female85yearsandover="B01001_049"),
                  state = c("MD"),
                  county = c("Allegany County", "Anne Arundel County", "Baltimore County", "Calvert County", "Carroll County", "Charles County", "Dorchester County", 
                             "Frederick County", "Howard County", "Montgomery County", "Prince George's County", "Somerset County", "St. Mary's County", "Washington County", 
                             "Wicomico County", "Worcester County"),
                  survey = "acs5",
                  output = "wide",
                  geometry = TRUE)

pop_pa_fem <- get_acs(geography = "block group", 
                  year = 2020,
                  variables = c(population = "B01001_001",
                                male= "B01001_002",
                                female = "B01001_026", 
                                FemaleUnder5years="B01001_027",
                                Female5to9years="B01001_028",
                                Female10to14years="B01001_029",
                                Female15to17years="B01001_030",
                                Female18and19years="B01001_031",
                                Female20years="B01001_032",
                                Female21years="B01001_033",
                                Female22to24years="B01001_034",
                                Female25to29years="B01001_035",
                                Female30to34years="B01001_036",
                                Female35to39years="B01001_037",
                                Female40to44years="B01001_038",
                                Female45to49years="B01001_039",
                                Female50to54years="B01001_040",
                                Female55to59years="B01001_041",
                                Female60and61years="B01001_042",
                                Female62to64years="B01001_043",
                                Female65and66years="B01001_044",
                                Female67to69years="B01001_045",
                                Female70to74years="B01001_046",
                                Female75to79years="B01001_047",
                                Female80to84years="B01001_048",
                                Female85yearsandover="B01001_049"),
                  state = c("PA"),
                  county = buffer_counties_state$NAMELSAD[buffer_counties_state$STUSPS=='PA'],
                  survey = "acs5",
                  output = "wide",
                  geometry = TRUE)

pop_ky_fem <- get_acs(geography = "block group", 
                  year = 2020,
                  variables = c(population = "B01001_001",
                                male= "B01001_002",
                                female = "B01001_026", 
                                FemaleUnder5years="B01001_027",
                                Female5to9years="B01001_028",
                                Female10to14years="B01001_029",
                                Female15to17years="B01001_030",
                                Female18and19years="B01001_031",
                                Female20years="B01001_032",
                                Female21years="B01001_033",
                                Female22to24years="B01001_034",
                                Female25to29years="B01001_035",
                                Female30to34years="B01001_036",
                                Female35to39years="B01001_037",
                                Female40to44years="B01001_038",
                                Female45to49years="B01001_039",
                                Female50to54years="B01001_040",
                                Female55to59years="B01001_041",
                                Female60and61years="B01001_042",
                                Female62to64years="B01001_043",
                                Female65and66years="B01001_044",
                                Female67to69years="B01001_045",
                                Female70to74years="B01001_046",
                                Female75to79years="B01001_047",
                                Female80to84years="B01001_048",
                                Female85yearsandover="B01001_049"),
                  state = c("KY"),
                  county = buffer_counties_state$NAMELSAD[buffer_counties_state$STUSPS=='KY'],
                  survey = "acs5",
                  output = "wide",
                  geometry = TRUE)

pop_nc_fem <- get_acs(geography = "block group", 
                  year = 2020,
                  variables = c(population = "B01001_001",
                                male= "B01001_002",
                                female = "B01001_026", 
                                FemaleUnder5years="B01001_027",
                                Female5to9years="B01001_028",
                                Female10to14years="B01001_029",
                                Female15to17years="B01001_030",
                                Female18and19years="B01001_031",
                                Female20years="B01001_032",
                                Female21years="B01001_033",
                                Female22to24years="B01001_034",
                                Female25to29years="B01001_035",
                                Female30to34years="B01001_036",
                                Female35to39years="B01001_037",
                                Female40to44years="B01001_038",
                                Female45to49years="B01001_039",
                                Female50to54years="B01001_040",
                                Female55to59years="B01001_041",
                                Female60and61years="B01001_042",
                                Female62to64years="B01001_043",
                                Female65and66years="B01001_044",
                                Female67to69years="B01001_045",
                                Female70to74years="B01001_046",
                                Female75to79years="B01001_047",
                                Female80to84years="B01001_048",
                                Female85yearsandover="B01001_049"),
                  state = c("NC"),
                  county = buffer_counties_state$NAMELSAD[buffer_counties_state$STUSPS=='NC'],
                  survey = "acs5",
                  output = "wide",
                  geometry = TRUE)

pop_tn_fem <- get_acs(geography = "block group", 
                  year = 2020,
                  variables = c(population = "B01001_001",
                                male= "B01001_002",
                                female = "B01001_026", 
                                FemaleUnder5years="B01001_027",
                                Female5to9years="B01001_028",
                                Female10to14years="B01001_029",
                                Female15to17years="B01001_030",
                                Female18and19years="B01001_031",
                                Female20years="B01001_032",
                                Female21years="B01001_033",
                                Female22to24years="B01001_034",
                                Female25to29years="B01001_035",
                                Female30to34years="B01001_036",
                                Female35to39years="B01001_037",
                                Female40to44years="B01001_038",
                                Female45to49years="B01001_039",
                                Female50to54years="B01001_040",
                                Female55to59years="B01001_041",
                                Female60and61years="B01001_042",
                                Female62to64years="B01001_043",
                                Female65and66years="B01001_044",
                                Female67to69years="B01001_045",
                                Female70to74years="B01001_046",
                                Female75to79years="B01001_047",
                                Female80to84years="B01001_048",
                                Female85yearsandover="B01001_049"),
                  state = c("TN"),
                  county = buffer_counties_state$NAMELSAD[buffer_counties_state$STUSPS=='TN'],
                  survey = "acs5",
                  output = "wide",
                  geometry = TRUE)

pop_wv_fem <- get_acs(geography = "block group", 
                  year = 2020,
                  variables = c(population = "B01001_001",
                                male= "B01001_002",
                                female = "B01001_026", 
                                FemaleUnder5years="B01001_027",
                                Female5to9years="B01001_028",
                                Female10to14years="B01001_029",
                                Female15to17years="B01001_030",
                                Female18and19years="B01001_031",
                                Female20years="B01001_032",
                                Female21years="B01001_033",
                                Female22to24years="B01001_034",
                                Female25to29years="B01001_035",
                                Female30to34years="B01001_036",
                                Female35to39years="B01001_037",
                                Female40to44years="B01001_038",
                                Female45to49years="B01001_039",
                                Female50to54years="B01001_040",
                                Female55to59years="B01001_041",
                                Female60and61years="B01001_042",
                                Female62to64years="B01001_043",
                                Female65and66years="B01001_044",
                                Female67to69years="B01001_045",
                                Female70to74years="B01001_046",
                                Female75to79years="B01001_047",
                                Female80to84years="B01001_048",
                                Female85yearsandover="B01001_049"),
                  state = c("WV"),
                  county = buffer_counties_state$NAMELSAD[buffer_counties_state$STUSPS=='WV'],
                  survey = "acs5",
                  output = "wide",
                  geometry = TRUE)

pop_de_fem <- get_acs(geography = "block group", 
                  year = 2020,
                  variables = c(population = "B01001_001",
                                male= "B01001_002",
                                female = "B01001_026", 
                                FemaleUnder5years="B01001_027",
                                Female5to9years="B01001_028",
                                Female10to14years="B01001_029",
                                Female15to17years="B01001_030",
                                Female18and19years="B01001_031",
                                Female20years="B01001_032",
                                Female21years="B01001_033",
                                Female22to24years="B01001_034",
                                Female25to29years="B01001_035",
                                Female30to34years="B01001_036",
                                Female35to39years="B01001_037",
                                Female40to44years="B01001_038",
                                Female45to49years="B01001_039",
                                Female50to54years="B01001_040",
                                Female55to59years="B01001_041",
                                Female60and61years="B01001_042",
                                Female62to64years="B01001_043",
                                Female65and66years="B01001_044",
                                Female67to69years="B01001_045",
                                Female70to74years="B01001_046",
                                Female75to79years="B01001_047",
                                Female80to84years="B01001_048",
                                Female85yearsandover="B01001_049"),
                  state = c("DE"),
                  county = buffer_counties_state$NAMELSAD[buffer_counties_state$STUSPS=='DE'],
                  survey = "acs5",
                  output = "wide",
                  geometry = TRUE)

#######

pop_va_buffer_fem <- rbind(pop_va_fem, pop_de_fem, pop_ky_fem, pop_md_fem, pop_nc_fem, pop_pa_fem, pop_tn_fem, pop_wv_fem)

va <- pop_va_buffer_fem %>% mutate( obgyn_pop= femaleE - FemaleUnder5yearsE -Female5to9yearsE - Female10to14yearsE ) %>% 
  select(GEOID, NAME, population=populationE, pop_fem= femaleE, obgyn_pop )

# data combined
#centroid and coordinates

data_combined <- data.frame(
  GEOID = va$GEOID,
  population = va$population,
  obgyn_pop = va$obgyn_pop,
st_coordinates(st_centroid(va$geometry))
)

data_combined <- data_combined %>% filter(!is.na(data_combined$X))

data_combined <- data_combined %>% filter(GEOID != 517000323001 )
data_combined <- data_combined %>% filter(GEOID != 517000323002 )
data_combined <- data_combined %>% filter(GEOID != 517000323003 )

```

# Travel time
```{r}
library(osrm)
options(osrm.server = "http://104.248.112.16:5001/", osrm.profile = "car")

if(!file.exists("traveltimes_exercise.csv")){
  traveltimes <- osrmTable(
    #src = data_combined[, c("GEOID", "X", "Y")],  #population-demand
    src = data_combined[, c( "X", "Y")],  #population-demand
    #dst = st_drop_geometry( provider[, c("ID", "lon", "lat")] )    #providers supply
    #dst =  provider[, c("ID", "lon", "lat")] #providers supply
    dst =  provider[, c( "lon", "lat")] #providers supply
  )$duration
  
  names(traveltimes) <- provider[, c("ID")]
  
  write.csv(
    # cbind(GEOID = rownames(traveltimes), as.data.frame(traveltimes)),
    cbind(GEOID = data_combined[, c("GEOID")], as.data.frame(traveltimes)),
    "traveltimes_exercise.csv", row.names = FALSE
  )
}

traveltimes <- read.csv("traveltimes_exercise.csv", row.names = 1)


```


# Define geography id - Geography-GEOID from initial file may be outdated
```{r}
library(tigris)
library(maps)
library(sf)
# add block geoids
# get US blocks shapefile

blocks_VA <- st_as_sf(block_groups(state= "VA", year=2020)) #, year=2010
blocks_DC <- st_as_sf(block_groups(state= "DC", year=2020)) #, year=2010
blocks_MD <- st_as_sf(block_groups(state= "MD", year=2020)) #, year=2010
blocks_NC <- st_as_sf(block_groups(state= "NC", year=2020)) #, year=2010
blocks_DE <- st_as_sf(block_groups(state= "DE", year=2020)) #, year=2010
blocks_TN <- st_as_sf(block_groups(state= "TN", year=2020)) #, year=2010
blocks_WV <- st_as_sf(block_groups(state= "WV", year=2020)) #, year=2010
blocks_KY <- st_as_sf(block_groups(state= "KY", year=2020)) #, year=2010
blocks_PA <- st_as_sf(block_groups(state= "PA", year=2020)) #, year=2010


blocks <- rbind( blocks_VA, blocks_DC, blocks_MD , blocks_NC , blocks_DE , blocks_TN , blocks_WV , blocks_KY , blocks_PA) 
# lon and lat to geo-points
geopts <- provider %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4269) #4326. initial: 4269
# indeces of bgs which contain a geopoint
inds <- st_within(geopts$geometry, blocks$geometry)
blk_list <- c()
for (i in inds){
  if (identical(blocks$NAME[i],character(0))){
    blk_list<- append(blk_list, NA)}
  else{
    blk_list <- append(blk_list, blocks$GEOID[i])}
}
provider['GEOID'] <- blk_list

```

# Add count of providers per geography using matching codes
```{r}

#providers w geoid
#provider$GEOID <- substr(provider$geoid_blk_new, 1, 12 ) 
#data_combined with geoid: ok
num_providers <- provider %>% group_by(GEOID) %>% summarise(prov_cnt = sum(doctors) )

#join providers to block groups
data_combined$GEOID <-  as.character( data_combined$GEOID)
data_combined <- data_combined %>% left_join(num_providers, by= "GEOID" )

sum(data_combined$prov_cnt, na.rm = TRUE)

```



# Add mean and median of 10 nearest drive times 
```{r}
#mean of 10 nearest
top_mean <- function(x) {  
   mean(head(sort(x ), 10) ) }
#median of 10 nearest
top_median <- function(x) {  
   median(head(sort(x ), 10) ) }
#apply rowwise
traveltimes_near <- data.frame(near_10_mean= base::apply(traveltimes, 1, top_mean), 
                               near_10_median= base::apply(traveltimes, 1, top_median)) 
#rownames_to_column(traveltimes_near, var = "GEOID")
traveltimes_near$GEOID <- row.names(traveltimes_near) 
#join mean median traveltimes to geographies
data_combined <- data_combined %>% left_join(traveltimes_near, by= "GEOID")
#data_combined <- st_drop_geometry(data_combined)

```

# Prepara data to save
```{r}
#raw traveltimes: traveltimes matrix already estimated and with colnames arranged 
traveltimes <- read.csv("traveltimes_exercise.csv", row.names = 1)
#population: always recheck relevant population: ie. for pediatrics: pop 0-17 years
population <- data_combined %>% select(GEOID, X, Y, obgyn_pop, prov_cnt, near_10_mean, near_10_median)
# realign travel times
traveltimes <- traveltimes[as.character(population$GEOID[1:nrow(traveltimes) ]), provider$ID]

```
# Save files to estimate catchment areas
```{r}
write.csv(provider[, c("ID", "address", "lat", "lon", "doctors")], "provider_20.csv", row.names = FALSE)
write.csv(cbind(GEOID = rownames(traveltimes), traveltimes), "traveltimes_trimmed_20.csv", row.names = FALSE)
write.csv(population, "population.csv", row.names = FALSE)
```



