---
title: "catchment_scores"
author: "--"
date: "2/4/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# libraries
```{r}
library(catchment)
library(sf)
library(leaflet)
library(dplyr)
library(catchment)
```
# start new
```{r}
 rm(list=ls())
```
#working directory
```{r}
setwd("~/VDH/Floating_Catchment_Areas/dmv/1. obgyn_dmv_model/")
```
# read data
```{r}
population <- read.csv("population.csv")
provider <- read.csv("provider.csv")
rownames(provider) <- provider$ID
traveltimes <- read.csv("traveltimes_trimmed.csv", row.names = 1)
```
# prepare map
```{r}
blockgroup_shapes <- st_transform(read_sf("blockgroups.geojson", as_tibble = FALSE), 4326)
rownames(blockgroup_shapes) <- blockgroup_shapes$GEOID
blockgroup_shapes <- blockgroup_shapes[as.character(population$GEOID),]
##new
shape_counties <- tigris::counties(state = c("VA", "MD", "DC"), class = "sf")
shape_counties_dmv <- shape_counties %>% filter(GEOID == 51013 | GEOID ==51059 | GEOID == 51600 | GEOID == 51107 | GEOID == 51610 | GEOID == 51683 | 
                                                  GEOID == 51685 | GEOID == 51153 | GEOID == 51510 | GEOID == 11001 | GEOID ==  24021 | GEOID == 24031 | GEOID == 24033 | GEOID == 24017) 
#format for counties
shape_counties_dmv <- st_transform(shape_counties_dmv, 4326 )

#create map
map <- leaflet(blockgroup_shapes, options = leafletOptions(attributionControl = FALSE)) %>%
  addTiles() %>%
  addScaleBar("bottomleft") %>%
  addMapPane("lines", zIndex = 410) %>%
  addMapPane("points", zIndex = 411) %>%
  addPolygons(
    fillColor = colorNumeric("RdYlBu", population$population)(population$population),
    fillOpacity = 1, stroke = FALSE, group = "Population", label = population$pediatric_pop
  ) %>%
  hideGroup("Population") %>%
  addLayersControl(
    position = "topleft", overlayGroups = c("Pediatrics", "Population", "Access")
  ) %>%
  addCircles(
    data = provider, color = "blue", lng = ~lon, lat = ~lat, weight = 3,
    label = ~ paste0("ID: ", ID), group = "ObGyns", options = pathOptions(pane = "points")
  )
```
# calculate and display catchment areas
## 2-step
```{r}
population$providers_2sfca <- catchment_ratio(
  population, provider, traveltimes, 30,
  consumers_value = "female_over_14", providers_id = "ID", providers_value = "doctors", verbose = TRUE
) * 1000

pal <- colorBin("RdYlBu", population$providers_2sfca)
plot_2sfca <-  map %>%
  addControl("Pediatrics Per 1,000 People (2-Step Floating Catchment Area)", "topright") %>%
  addLegend("bottomright", pal, population$providers_2sfca, opacity = .7) %>%
  addPolygons(
    fillColor = pal(population$providers_2sfca), fillOpacity = .7, weight = 1, color = "#000",
    highlightOptions = highlightOptions(color = "#fff"), group = "Access",
    label = paste0(
      "GEOID: ", population$GEOID, ", Population Ages 0-17: ", population$population,
      ", Per 1k People: ", round(population$providers_2sfca, 4), ", In Region: ",
      round(population$population * population$providers_2sfca / 1000, 4)
    )
  ) %>%
  addPolylines(data = shape_counties_dmv, color = "black", opacity = 1, weight = 2.5) 
```
## enhanced 2-step
```{r}
weight <- list(c(60, .042), c(30, .377), c(20, .704), c(10, .962))
population$providers_e2sfca <- catchment_ratio(
  population, provider, traveltimes, weight,
  consumers_value = "female_over_14", providers_id = "ID", providers_value = "doctors", verbose = TRUE
) * 1000

pal <- colorBin("RdYlBu", population$providers_e2sfca)
plot_e2sfca <- map %>%
  addControl("Pediatrics Per 1,000 People (enhanced 2-Step Floating Catchment Area)", "topright") %>%
  addLegend("bottomright", pal, population$providers_e2sfca, opacity = .7) %>%
  addPolygons(
    fillColor = pal(population$providers_e2sfca), fillOpacity = .7, weight = 1, color = "#000",
    highlightOptions = highlightOptions(color = "#fff"), group = "Access",
    label = paste0(
      "GEOID: ", population$GEOID, ", Population Ages 0-17: ", population$population,
      ", Per 1k People: ", round(population$providers_e2sfca, 4), ", In Region: ",
      round(population$population * population$providers_e2sfca / 1000, 4)
    )
  ) %>% addPolylines(data = shape_counties_dmv, color = "black", opacity = 1, weight = 2.5) 

```
## 3-step
```{r}
population$providers_3sfca <- catchment_ratio(
  population, provider, traveltimes, "gaussian", normalize_weight = TRUE, scale = 20,
  consumers_value = "female_over_14", providers_id = "ID", providers_value = "doctors", verbose = TRUE
) * 1000

pal <- colorBin("RdYlBu", population$providers_3sfca)
plot_3sfca <- map %>%
  addControl("Pediatrics Per 1,000 People (3-Step Floating Catchment Area)", "topright") %>%
  addLegend("bottomright", pal, population$providers_3sfca, opacity = .7) %>%
  addPolygons(
    fillColor = pal(population$providers_3sfca), fillOpacity = .7, weight = 1, color = "#000",
    highlightOptions = highlightOptions(color = "#fff"), group = "Access",
    label = paste0(
      "GEOID: ", population$GEOID, ", Population Ages 0-17: ", population$population,
      ", Per 1k People: ", round(population$providers_3sfca, 4), ", In Region: ",
      round(population$population * population$providers_3sfca / 1000, 4)
    )
  ) %>% addPolylines(data = shape_counties_dmv, color = "black", opacity = 1, weight = 2.5) 
```
#plot maps
```{r}
plot_2sfca
plot_e2sfca
plot_3sfca
```


#reshape wide-long
```{r}
######

#DEFINE relevant population and names
dmv_bg_2021_acccess_scores_wide <- population   %>% dplyr::select( geoid="GEOID", 
                                                            obgyn_pop_cnt ="female_over_14",
                                                            obgyn_cnt = "doctors",
                                                            obgyn_near_10_mean = "near_10_mean",
                                                            obgyn_near_10_median = "near_10_median",
                                                            obgyn_2sfca = "providers_2sfca" ,
                                                            obgyn_e2sfca = "providers_e2sfca", 
                                                            obgyn_3sfca = "providers_3sfca" )
```

# long format block group
```{r}
dmv_bg_2021_acccess_scores_bg <- tidyr::gather(dmv_bg_2021_acccess_scores_wide, measure, value, 
                                            obgyn_pop_cnt,
                                            obgyn_cnt,
                                            obgyn_near_10_mean,
                                            obgyn_near_10_median,
                                            obgyn_2sfca, 
                                            obgyn_e2sfca, 
                                            obgyn_3sfca)
#define year
dmv_bg_2021_acccess_scores_bg$year <- 2021
#define measure type
dmv_bg_2021_acccess_scores_bg$measure_type <- as.factor(ifelse(dmv_bg_2021_acccess_scores_bg$measure=="obgyn_2sfca" |
                                                            dmv_bg_2021_acccess_scores_bg$measure=="obgyn_e2sfca" |
                                                              dmv_bg_2021_acccess_scores_bg$measure=="obgyn_3sfca", 'index','count') )
```

#include name: geographic name for the smallest geographic unit: block group in this case
```{r}
#a) connect
con <- RPostgreSQL::dbConnect(drv = RPostgreSQL::PostgreSQL(),
                              dbname = "sdad",
                              host = "postgis1",
                              port = 5432,
                              user = Sys.getenv(x = "DB_USR"),
                              password = Sys.getenv(x = "DB_PWD"))

#b) query
name_geo <- sf::st_read(con, 
  query= "
SELECT  *
FROM dc_geographies.ncr_cttrbg_tiger_2010_2020_geo_names "
)

#join region_name + region_type
dmv_bg_2021_acccess_scores_bg$geoid <- as.character(dmv_bg_2021_acccess_scores_bg$geoid)
dmv_bg_2021_acccess_scores_bg <- dmv_bg_2021_acccess_scores_bg %>% left_join(name_geo, by="geoid")
#order
dmv_bg_2021_acccess_scores_bg <- dmv_bg_2021_acccess_scores_bg %>% dplyr::select("geoid","region_type","region_name", "measure", "value", "year", "measure_type")
######
```

# AGGREGATION tract
```{r}
#generate id tract 
#the relevant file is WIDE
dmv_bg_2021_acccess_scores_wide_aggreg <- dmv_bg_2021_acccess_scores_wide  %>% mutate(tract_geoid = substr(geoid,1,11))
#add region name - tract
# dmv_bg_2021_acccess_scores_wide_aggreg <- dmv_bg_2021_acccess_scores_wide_aggreg %>% left_join(name_geo, by= c("tract_geoid"= "geoid") )

## measures tract
dmv_bg_2021_acccess_scores_tract_wide <- dmv_bg_2021_acccess_scores_wide_aggreg %>% group_by(tract_geoid) %>% summarise(
  obgyn_pop = sum(obgyn_pop_cnt, na.rm = TRUE), 
  obgyn_cnt = sum(obgyn_cnt, na.rm = TRUE), 
  obgyn_near_10_mean = weighted.mean(obgyn_near_10_mean, obgyn_pop_cnt, na.rm = TRUE), 
  obgyn_near_10_median = weighted.mean(obgyn_near_10_median, obgyn_pop_cnt, na.rm = TRUE), 
  obgyn_2sfca = weighted.mean(obgyn_2sfca, obgyn_pop_cnt, na.rm = TRUE), 
  obgyn_e2sfca = weighted.mean(obgyn_e2sfca, obgyn_pop_cnt, na.rm = TRUE),
  obgyn_3sfca = weighted.mean(obgyn_3sfca, obgyn_pop_cnt, na.rm = TRUE),
  )

#write the name of population again, the name was change to avoid conflict when calculating weighted mean
names(dmv_bg_2021_acccess_scores_tract_wide)[2] <- 'obgyn_pop_cnt'
               
```

# long format tract
```{r}
dmv_bg_2021_acccess_scores_tract <- tidyr::gather(dmv_bg_2021_acccess_scores_tract_wide, measure, value, 
                                            obgyn_pop_cnt,
                                            obgyn_cnt,
                                            obgyn_near_10_mean,
                                            obgyn_near_10_median,
                                            obgyn_2sfca, 
                                            obgyn_e2sfca, 
                                            obgyn_3sfca)
#define year
dmv_bg_2021_acccess_scores_tract$year <- 2021
#define measure type
dmv_bg_2021_acccess_scores_tract$measure_type <- as.factor(ifelse(dmv_bg_2021_acccess_scores_tract$measure=="obgyn_2sfca" |
                                                            dmv_bg_2021_acccess_scores_tract$measure=="obgyn_e2sfca" |
                                                              dmv_bg_2021_acccess_scores_tract$measure=="obgyn_3sfca", 'index','count') )
```

#include name: geographic name for geographic unit: tract in this case
```{r}
#join region_name + region_type
dmv_bg_2021_acccess_scores_tract$tract_geoid <- as.character(dmv_bg_2021_acccess_scores_tract$tract_geoid)
dmv_bg_2021_acccess_scores_tract <- dmv_bg_2021_acccess_scores_tract %>% left_join(name_geo, by= c("tract_geoid" = "geoid") )
#order
dmv_bg_2021_acccess_scores_tract <- dmv_bg_2021_acccess_scores_tract %>% dplyr::select(geoid="tract_geoid","region_type","region_name", "measure", "value", "year", "measure_type")
```


# AGGREGATION county
```{r}
#generate id tract 
#the relevant file is WIDE
dmv_bg_2021_acccess_scores_wide_aggreg <- dmv_bg_2021_acccess_scores_wide  %>% mutate(county_geoid = substr(geoid,1,5))

## measures tract
dmv_bg_2021_acccess_scores_county_wide <- dmv_bg_2021_acccess_scores_wide_aggreg %>% group_by(county_geoid) %>% summarise(
  obgyn_pop = sum(obgyn_pop_cnt, na.rm = TRUE), 
  obgyn_cnt = sum(obgyn_cnt, na.rm = TRUE), 
  obgyn_near_10_mean = weighted.mean(obgyn_near_10_mean, obgyn_pop_cnt, na.rm = TRUE), 
  obgyn_near_10_median = weighted.mean(obgyn_near_10_median, obgyn_pop_cnt, na.rm = TRUE), 
  obgyn_2sfca = weighted.mean(obgyn_2sfca, obgyn_pop_cnt, na.rm = TRUE), 
  obgyn_e2sfca = weighted.mean(obgyn_e2sfca, obgyn_pop_cnt, na.rm = TRUE),
  obgyn_3sfca = weighted.mean(obgyn_3sfca, obgyn_pop_cnt, na.rm = TRUE),
  )

#write the name of population again, the name was change to avoid conflict when calculating weighted mean
names(dmv_bg_2021_acccess_scores_county_wide)[2] <- 'obgyn_pop_cnt'
               
```

# long format tract
```{r}
dmv_bg_2021_acccess_scores_county <- tidyr::gather(dmv_bg_2021_acccess_scores_county_wide, measure, value, 
                                            obgyn_pop_cnt,
                                            obgyn_cnt,
                                            obgyn_near_10_mean,
                                            obgyn_near_10_median,
                                            obgyn_2sfca, 
                                            obgyn_e2sfca, 
                                            obgyn_3sfca)
#define year
dmv_bg_2021_acccess_scores_county$year <- 2021
#define measure type
dmv_bg_2021_acccess_scores_county$measure_type <- as.factor(ifelse(dmv_bg_2021_acccess_scores_county$measure=="obgyn_2sfca" |
                                                            dmv_bg_2021_acccess_scores_county$measure=="obgyn_e2sfca" |
                                                              dmv_bg_2021_acccess_scores_county$measure=="obgyn_3sfca", 'index','count') )
```

#include name: geographic name for geographic unit: county in this case
```{r}
#join region_name + region_type
dmv_bg_2021_acccess_scores_county$tract_geoid <- as.character(dmv_bg_2021_acccess_scores_county$county_geoid)
dmv_bg_2021_acccess_scores_county <- dmv_bg_2021_acccess_scores_county %>% left_join(name_geo, by= c("county_geoid" = "geoid") )
#order
dmv_bg_2021_acccess_scores_county <- dmv_bg_2021_acccess_scores_county %>% dplyr::select(geoid="county_geoid","region_type","region_name", "measure", "value", "year", "measure_type")
```

# stack all the measures for BLOCK GROUP + TRACT + COUNTY: include NAME
```{r}
dmv_bg_2021_acccess_scores_obgyn <- rbind(dmv_bg_2021_acccess_scores_bg, 
                                    dmv_bg_2021_acccess_scores_tract,
                                    dmv_bg_2021_acccess_scores_county)
```

```{r}
#SAVE with specific name
# write.csv(dmv_bg_2021_acccess_scores_obgyn, "~/VDH/Floating_Catchment_Areas/dmv/1. obgyn_dmv_model/dmv_cttrbg_webmd_2021_access_scores_obgyn.csv", row.names = FALSE )
readr::write_csv(dmv_bg_2021_acccess_scores_obgyn, xzfile("ncr_cttrbg_webmd_2021_access_scores_obgyn.csv.xz", compression = 9))

# #a) connect
# con <- RPostgreSQL::dbConnect(drv = RPostgreSQL::PostgreSQL(),
#                                dbname = "sdad",
#                                host = "postgis1",
#                                port = 5432,
#                                user = Sys.getenv(x = "DB_USR"),
#                                password = Sys.getenv(x = "DB_PWD"))
# 
# #b) save the file in scheme dc_working in pgadmin: CHANGE name
# sf::st_write(
#   dmv_bg_2021_acccess_scores_obgyn, con,
#   c("dc_health_behavior_diet", "dmv_cttrbg_webmd_2021_access_scores_obgyn"),
#   table_owner = "data_commons",
#   row.names = FALSE)
# 
# #d) Disconnect
# RPostgreSQL::dbDisconnect(con)
```

