---
title: "Create Datafile"
author: "Aaron Schroeder"
date: "11/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Get dataset DOI

```{r}
dataset_doi <- readLines(here::here("data/dataset_doi"))[1]
```

## Code you used to create dataset - LIST BELOW

dc_digital_communications.va_tr_broadband_now_2021_internet_package_price_update,
dc_digital_communications.va_ct_broadband_now_2021_internet_package_price,
dc_digital_communications.va_hd_broadband_now_2021_internet_package_price: DMV_Internet_Prices.RMD

dc_digital_communications.va_tr_broadband_now_2021_internet_package_price_100_20,
dc_digital_communications.va_ct_broadband_now_2021_internet_package_price_100_20,
dc_digital_communications.va_hd_broadband_now_2021_internet_package_price_100_20: Price_Code_100_20.RMD


```{r data}

# Ex database_table_name: "va_hd_vdoe_2019_2021_3rd_grade_mean_median_read_score"
table_names = c("va_tr_broadband_now_2021_internet_package_price_update",
                "va_ct_broadband_now_2021_internet_package_price",
                "va_hd_broadband_now_2021_internet_package_price",
                "va_tr_broadband_now_2021_internet_package_price_100_20",
                "va_ct_broadband_now_2021_internet_package_price_100_20",
                "va_hd_broadband_now_2021_internet_package_price_100_20")
    
con <- get_db_conn()

dat_file_name <- table_names[1]
db_schema_name <- "dc_digital_communications"  # Ex database schema_name: dc_education_training
all_geos_dat <- DBI::dbReadTable(con, c(db_schema_name, dat_file_name))  

for (i in 2:length(table_names))
{ 
  dat_file_name <- table_names[i]
  dat <- DBI::dbReadTable(con, c(db_schema_name, dat_file_name)) 

  all_geos_dat <- rbind(all_geos_dat, dat)  
}

DBI::dbDisconnect(con)

all_geos_file_name <- "va_hdcttr_2021_broadband_now_internet_package_price"  #Ex: "va_vdoe_2019_2021_3rd_grade_mean_median_read_score"
assign(all_geos_file_name, all_geos_dat)
```

## Write to compressed file

```{r compress}
dat_file_path <- here::here(paste0("data/", all_geos_file_name, ".csv.xz"))

readr::write_csv(get(all_geos_file_name),
                 xzfile(dat_file_path, compression = 9))
```

## Upload Files to Dataverse Dataset

```{r upload}
dataverse::add_dataset_file(
  file = dat_file_path,
  dataset = dataset_doi,
  key = Sys.getenv("DATAVERSE_KEY"),
  server   = Sys.getenv("DATAVERSE_SERVER"),
  description = all_geos_file_name
)
```
