---
title: "Internet_Price"
output: html_document
---

# Idea here - from Corelogic, get a couple of addresses from each block group (or nearby)
# Install the dc.utils, which contain code to access the database
```{r}
#install.packages("remotes")
library(devtools)
install_github("uva-bi-sdad/sdc.utils")
library(dc.utils)

```

# List all available functions in util
```{r}
# lsf.str("package:utils")
exists('get_db_conn', where='package:dc.utils', mode='function')
exists('st_read', where='package:dc.utils', mode='function')
```

# Print the current working directory, and check if the credentials exist
# ```{r}
# list.files(getwd())
# file.exists("~/.Renviron")
# readRenviron("~/.Renviron")
# Sys.getenv("db_nam")
# ```

# Assuming you don't have access to the function

<!-- ```{r} -->
<!-- get_db_conn <- -->
<!--   function(db_name = Sys.getenv("db_nam"), -->
<!--            db_host = Sys.getenv("db_hst"), -->
<!--            db_port = Sys.getenv("db_prt"), -->
<!--            db_user = Sys.getenv("db_usr"), -->
<!--            db_pass = Sys.getenv("db_pwd")) { -->
<!--     RPostgreSQL::dbConnect( -->
<!--       drv = RPostgreSQL::PostgreSQL(), -->
<!--       dbname = db_name, -->
<!--       host = db_host, -->
<!--       port = db_port, -->
<!--       user = db_user, -->
<!--       password = db_pass -->
<!--     ) -->
<!--   } -->
<!-- ``` -->



# load packages
```{r}
library(sf)
library(tidyverse)
library(tmap)
library(tmaptools)
library(tigris)
library(tidycensus)
library(rmapshaper)
library(matrixStats)
library(SpatialAcc)
library(reticulate)
library(dplyr)
library(tidygeocoder)
library(readxl)
library(DBI)
library(sf)
library(mapview)
library(progress)
options(tigris_use_cache = TRUE)
```

# download county-level data, and plot to verify

```{r}
usa <- get_acs(geography = "county",
                 year = 2019,
                 variables = c(median_household_income = "B19013_001"),
                 survey = "acs5",
                 output = "wide",
                 geometry = T) 


shifted_usa <- usa%>%shift_geometry() # need gemoetry for st_centroid

plot(shifted_usa["median_household_incomeE"])
```

# Getting the centroid for each of the geometries (see https://stackoverflow.com/questions/54352407/tidycensus-package-can-you-get-latitudes-and-longitudes-with-census-blocks)
```{r}
count_centroid <- shifted_usa %>% 
  #st_transform(2273) %>% # convert to projected coord system for better centroid
  st_centroid()

count_centroid

ggplot(shifted_usa) +
  geom_sf() +
  geom_sf(data = count_centroid, size=0.25) +
  theme_void()
ggsave("../../docs/usa_county_centroid.pdf")
```

# Add n random points per county as backup
# https://stackoverflow.com/questions/70072417/generating-random-points-within-sub-geometries-using-st-sample-from-the-sf-packa
```{r}
# Good idea to test on smaller geometry first
county <- get_acs(geography = "county",
                 year = 2019,
                 state = "NY",
                 variables = c(median_household_income = "B19013_001"),
                 survey = "acs5",
                 output = "wide",
                 geometry = T) 


county

sample_points <- county %>% 
  st_sample(size = c(3,3 ))

sample_points
ggplot(county) +
  geom_sf() +
  geom_sf(data = sample_points, size=0.25) +
  theme_void()
ggsave("../../docs/county_sample.pdf")
```

```{r}
sample_points <- usa %>% 
  st_sample(size = c(3,3 ))

sample_points
```

```{r}
count_centroid <- usa %>% 
  st_centroid()

head(count_centroid)
library(mapview)
mapview(count_centroid)
```


```{r}
mapview(sample_points)
```


# Get the lat lon of the positions
```{r}
count_centroid
count_centroid <- cbind(usa, st_coordinates(count_centroid))
# count_centroid <- count_centroid %>% extract("geometry", c('lat', 'lon'), '\\((.*), (.*)\\)', convert = TRUE)
names(count_centroid)[names(count_centroid) == 'X'] <- 'lon'
names(count_centroid)[names(count_centroid) == 'Y'] <- 'lat'
colnames(count_centroid)
```


```{r}
# set.seed(1) 

sample_points

# duplicate each row 3 times to match the n number of samples
usa_3 <- usa %>% slice(rep(1:n(), each = 3))

samp <- cbind(usa_3, st_coordinates(sample_points))
samp

# count_centroid <- count_centroid %>% extract("geometry", c('lat', 'lon'), '\\((.*), (.*)\\)', convert = TRUE)
names(samp)[names(samp) == 'X'] <- 'lon'
names(samp)[names(samp) == 'Y'] <- 'lat'
colnames(samp)
samp
```

```{r}
colnames(count_centroid)
colnames(samp)


combined <- rbind(count_centroid, samp)
combined
```


<!-- # Not all of the addresses returned a price point, so instead we should retrieve N points within each county to reverse geo code (https://stackoverflow.com/questions/52248394/get-census-tract-from-lat-lon-using-tigris) -->

# Reverse geocode based on the lat long of the centroid. Reference: https://walker-data.com/census-r/census-geographic-data-and-applications-in-r.html
```{r}
# centroid_address_path <- "../../data/usa_reverse_geocode_county_centroid.csv"


# if (file.exists(centroid_address_path)){
#   print("Centroid file exists, loading...")
#   reverse = read.csv(centroid_address_path)
# }else{ # Took me 3230.6 seconds
#   
  
increment <- 10
n_iter <- nrow(combined)%/% increment # integer divide

pb <- progress_bar$new(format = "(:spin) [:bar] :percent [Elapsed time: :elapsedfull || Estimated time remaining: :eta]",
                       total = n_iter,
                       complete = "=",   # Completion bar character
                       incomplete = "-", # Incomplete bar character
                       current = ">",    # Current bar character
                       clear = FALSE,    # If TRUE, clears the bar when finish
                       width = 100)      # Width of the progress bar

# nrow(combined)

for(i in seq(from=1, to=nrow(combined), by=increment)) { # reverse geocode has limit of 100
  pb$tick()
  frame <- combined[i:(i+(increment-1)),]
  frame
  reverse <- frame%>% 
    reverse_geocode(lat = "lat", long = "lon", method = 'osm', address = "address", full_results = FALSE)
  
  # print(reverse)
  
  rev <- subset(reverse, select=c('lat','lon','address','GEOID','NAME'))
  rev
  rev <- rev %>% 
       rename("geoid" = "GEOID",
              "name" = "NAME")
  rev
  # # frames[[(i%/% increment + 1)]] 
  write.csv(rev, paste('../../data/reverse_address/',i,'.csv', sep=' '), row.names=FALSE)  
}

```
```{r}
frames[[1]]
colnames(frames[[1]])
reversed <- bind_rows(frames)
reversed
```



```{r}

sample_address_path <- "../../data/usa_reverse_geocode_county.csv"
reverse['geoid'] <- combined$GEOID
reverse
write.csv(reverse, sample_address_path, row.names=FALSE)  
#}

colnames(reverse)
```

