---
title: "DMV_Internet_Prices"
output: html_document
---

# Important note - the naming should be improved here

# imports

```{r}
library(sf)
library(tidyverse)
library(tmap)
library(tmaptools)
library(tigris)
library(tidycensus)
library(rmapshaper)
library(matrixStats)
library(SpatialAcc)
library(reticulate)
library(dplyr)
library(tidygeocoder)
library(readxl)
```

# read in raw package data from step 2

```{r}
# read in data
con <- get_db_conn(db_pass = "rsu8zvrsu8zv")
va_broadband_now_prices <- st_read(con, query = "SELECT * FROM dc_digital_communications.va_bg_broadband_now_2021_all_internet_packages")
dc_broadband_now_prices <- st_read(con, query = "SELECT * FROM dc_digital_communications.dc_bg_broadband_now_2021_all_internet_packages")
md_broadband_now_prices <- st_read(con, query = "SELECT * FROM dc_digital_communications.md_bg_broadband_now_2021_all_internet_packages")
dmv_broadband_now_prices <- st_read(con, query = "SELECT * FROM dc_digital_communications.dmv_bg_broadband_now_2021_all_internet_packages")
dbDisconnect(con)

# add tract and county columns
md_broadband_now_prices <- md_broadband_now_prices %>%
  mutate(block_group = as.character(block_group),
         tract = substr(block_group, 1, 11),
         county = substr(block_group, 1, 5))
va_broadband_now_prices <- va_broadband_now_prices %>%
  mutate(block_group = as.character(block_group),
         tract = substr(block_group, 1, 11),
         county = substr(block_group, 1, 5))
dc_broadband_now_prices <- dc_broadband_now_prices %>%
  mutate(block_group = as.character(block_group),
         tract = substr(block_group, 1, 11),
         county = substr(block_group, 1, 5))
dmv_broadband_now_prices <- dmv_broadband_now_prices %>%
  mutate(block_group = as.character(block_group),
         tract = substr(block_group, 1, 11),
         county = substr(block_group, 1, 5))
```

# Get measures for VA, MD, and DC at tract level

```{r}
# VIRGINIA
va_100_min_prices <- va_broadband_now_prices %>%
  mutate(down_minus_100 = ifelse(download >= 100, download - 100, NA),) %>%
  drop_na(down_minus_100) %>%
  group_by(block_group) %>%
  slice_max(order_by = -down_minus_100, n = 1) %>%
  slice_max(order_by = -price, n = 1)  %>%
  group_by(tract) %>%
  summarize(min_price_100 = median(price, na.rm = T)) %>%
  mutate(tract = as.character(tract))
va_100_min_prices.2 <- va_broadband_now_prices %>%
  mutate(down_minus_100 = ifelse(download >= 100, download - 100, NA)) %>%
  drop_na(down_minus_100) %>%
  group_by(block_group) %>%
  slice_max(order_by = -price, n = 1)  %>%
  group_by(tract) %>%
  summarize(min_price_100_plus = median(price, na.rm = T)) %>%
  mutate(tract = as.character(tract)) %>%
  full_join(va_100_min_prices, by = "tract")
va_100_med_and_min_prices <- va_broadband_now_prices %>%
  mutate(down_minus_100 = ifelse(download >= 100, download - 100, NA)) %>%
  drop_na(down_minus_100) %>%
  group_by(block_group) %>%
  slice_max(order_by = -down_minus_100, n = 1) %>%
  group_by(tract) %>%
  summarize(med_price_100 = median(price, na.rm = T)) %>%
  mutate(tract = as.character(tract)) %>%
  full_join(va_100_min_prices.2, by = "tract")
va_min_prices <- va_broadband_now_prices %>% # find minimum price for 25 in each block group, then median price among those in tract
  mutate(down_minus_25 = ifelse(download >= 25, download - 25, NA)) %>%
  drop_na(down_minus_25) %>%
  mutate(down_minus_100 = ifelse(download <= 100, 100 - download, NA)) %>%
  drop_na(down_minus_100) %>%
  group_by(block_group) %>%
  slice_max(order_by = -price, n = 1)  %>%
  group_by(tract) %>%
  summarize(min_price_25 = median(price, na.rm = T)) %>%
  mutate(tract = as.character(tract)) %>%
  full_join(va_100_med_and_min_prices, by = "tract")
va_med_and_min_prices <- va_broadband_now_prices %>%
  mutate(down_minus_25 = ifelse(download >= 25, download - 25, NA)) %>%
  drop_na(down_minus_25) %>%
  mutate(down_minus_100 = ifelse(download <= 100, 100 - download, NA)) %>%
  drop_na(down_minus_100) %>%
  group_by(block_group) %>%
  summarize(price = median(price, na.rm = T), tract = tract) %>%
  group_by(tract) %>%
  summarize(med_price_25 = median(price, na.rm = T)) %>%
  mutate(tract = as.character(tract)) %>%
  full_join(va_min_prices, by = "tract")

# MARYLAND
md_100_min_prices <- md_broadband_now_prices %>%
  mutate(down_minus_100 = ifelse(download >= 100, download - 100, NA)) %>%
  drop_na(down_minus_100) %>%
  group_by(block_group) %>%
  slice_max(order_by = -down_minus_100, n = 1) %>%
  slice_max(order_by = -price, n = 1)  %>%
  group_by(tract) %>%
  summarize(min_price_100 = median(price, na.rm = T)) %>%
  mutate(tract = as.character(tract))
md_100_min_prices.2 <- md_broadband_now_prices %>%
  mutate(down_minus_100 = ifelse(download >= 100, download - 100, NA)) %>%
  drop_na(down_minus_100) %>%
  group_by(block_group) %>%
  slice_max(order_by = -price, n = 1)  %>%
  group_by(tract) %>%
  summarize(min_price_100_plus = median(price, na.rm = T)) %>%
  mutate(tract = as.character(tract)) %>%
  full_join(md_100_min_prices, by = "tract")
md_100_med_and_min_prices <- md_broadband_now_prices %>%
  mutate(down_minus_100 = ifelse(download >= 100, download - 100, NA)) %>%
  drop_na(down_minus_100) %>%
  group_by(block_group) %>%
  slice_max(order_by = -down_minus_100, n = 1) %>%
  group_by(tract) %>%
  summarize(med_price_100 = median(price, na.rm = T)) %>%
  mutate(tract = as.character(tract)) %>%
  full_join(md_100_min_prices.2, by = "tract")
md_min_prices <- md_broadband_now_prices %>% # find minimum price for 25 in each block group, then median price among those in tract
  mutate(down_minus_25 = ifelse(download >= 25, download - 25, NA)) %>%
  drop_na(down_minus_25) %>%
  mutate(down_minus_100 = ifelse(download <= 100, 100 - download, NA)) %>%
  drop_na(down_minus_100) %>%
  group_by(block_group) %>%
  slice_max(order_by = -price, n = 1)  %>%
  group_by(tract) %>%
  summarize(min_price_25 = median(price, na.rm = T)) %>%
  mutate(tract = as.character(tract)) %>%
  full_join(md_100_med_and_min_prices, by = "tract")
md_med_and_min_prices <- md_broadband_now_prices %>%
  mutate(down_minus_25 = ifelse(download >= 25, download - 25, NA)) %>%
  drop_na(down_minus_25) %>%
  mutate(down_minus_100 = ifelse(download <= 100, 100 - download, NA)) %>%
  drop_na(down_minus_100) %>%
  group_by(block_group) %>%
  summarize(price = median(price, na.rm = T), tract = tract) %>%
  group_by(tract) %>%
  summarize(med_price_25 = median(price, na.rm = T)) %>%
  mutate(tract = as.character(tract)) %>%
  full_join(md_min_prices, by = "tract")

# DC
dc_100_min_prices <- dc_broadband_now_prices %>%
  mutate(down_minus_100 = ifelse(download >= 100, download - 100, NA)) %>%
  drop_na(down_minus_100) %>%
  group_by(block_group) %>%
  slice_max(order_by = -down_minus_100, n = 1) %>%
  slice_max(order_by = -price, n = 1)  %>%
  group_by(tract) %>%
  summarize(min_price_100 = median(price, na.rm = T)) %>%
  mutate(tract = as.character(tract))
dc_100_min_prices.2 <- dc_broadband_now_prices %>%
  mutate(down_minus_100 = ifelse(download >= 100, download - 100, NA)) %>%
  drop_na(down_minus_100) %>%
  group_by(block_group) %>%
  slice_max(order_by = -price, n = 1)  %>%
  group_by(tract) %>%
  summarize(min_price_100_plus = median(price, na.rm = T)) %>%
  mutate(tract = as.character(tract)) %>%
  full_join(dc_100_min_prices, by = "tract")
dc_100_med_and_min_prices <- dc_broadband_now_prices %>%
  mutate(down_minus_100 = ifelse(download >= 100, download - 100, NA)) %>%
  drop_na(down_minus_100) %>%
  group_by(block_group) %>%
  slice_max(order_by = -down_minus_100, n = 1) %>%
  group_by(tract) %>%
  summarize(med_price_100 = median(price, na.rm = T)) %>%
  mutate(tract = as.character(tract)) %>%
  full_join(dc_100_min_prices.2, by = "tract")
dc_min_prices <- dc_broadband_now_prices %>% # find minimum price for 25 in each block group, then median price among those in tract
  mutate(down_minus_25 = ifelse(download >= 25, download - 25, NA)) %>%
  drop_na(down_minus_25) %>%
  mutate(down_minus_100 = ifelse(download <= 100, 100 - download, NA)) %>%
  drop_na(down_minus_100) %>%
  group_by(block_group) %>%
  slice_max(order_by = -price, n = 1)  %>%
  group_by(tract) %>%
  summarize(min_price_25 = median(price, na.rm = T)) %>%
  mutate(tract = as.character(tract)) %>%
  full_join(dc_100_med_and_min_prices, by = "tract")
dc_med_and_min_prices <- dc_broadband_now_prices %>%
  mutate(down_minus_25 = ifelse(download >= 25, download - 25, NA)) %>%
  drop_na(down_minus_25) %>%
  mutate(down_minus_100 = ifelse(download <= 100, 100 - download, NA)) %>%
  drop_na(down_minus_100) %>%
  group_by(block_group) %>%
  summarize(price = median(price, na.rm = T), tract = tract) %>%
  group_by(tract) %>%
  summarize(med_price_25 = median(price, na.rm = T)) %>%
  mutate(tract = as.character(tract)) %>%
  full_join(dc_min_prices, by = "tract")

# DMV
dmv_100_min_prices <- dmv_broadband_now_prices %>%
  mutate(down_minus_100 = ifelse(download >= 100, download - 100, NA)) %>%
  drop_na(down_minus_100) %>%
  group_by(block_group) %>%
  slice_max(order_by = -down_minus_100, n = 1) %>%
  slice_max(order_by = -price, n = 1)  %>%
  group_by(tract) %>%
  summarize(min_price_100 = median(price, na.rm = T)) %>%
  mutate(tract = as.character(tract))
dmv_100_min_prices.2 <- dmv_broadband_now_prices %>%
  mutate(down_minus_100 = ifelse(download >= 100, download - 100, NA)) %>%
  drop_na(down_minus_100) %>%
  group_by(block_group) %>%
  slice_max(order_by = -price, n = 1)  %>%
  group_by(tract) %>%
  summarize(min_price_100_plus = median(price, na.rm = T)) %>%
  mutate(tract = as.character(tract)) %>%
  full_join(dmv_100_min_prices, by = "tract")
dmv_100_med_and_min_prices <- dmv_broadband_now_prices %>%
  mutate(down_minus_100 = ifelse(download >= 100, download - 100, NA)) %>%
  drop_na(down_minus_100) %>%
  group_by(block_group) %>%
  slice_max(order_by = -down_minus_100, n = 1) %>%
  group_by(tract) %>%
  summarize(med_price_100 = median(price, na.rm = T)) %>%
  mutate(tract = as.character(tract)) %>%
  full_join(dmv_100_min_prices.2, by = "tract")
dmv_min_prices <- dmv_broadband_now_prices %>% # find minimum price for 25 in each block group, then median price among those in tract
  mutate(down_minus_25 = ifelse(download >= 25, download - 25, NA)) %>%
  drop_na(down_minus_25) %>%
  mutate(down_minus_100 = ifelse(download <= 100, 100 - download, NA)) %>%
  drop_na(down_minus_100) %>%
  group_by(block_group) %>%
  slice_max(order_by = -price, n = 1)  %>%
  group_by(tract) %>%
  summarize(min_price_25 = median(price, na.rm = T)) %>%
  mutate(tract = as.character(tract)) %>%
  full_join(dmv_100_med_and_min_prices, by = "tract")
dmv_med_and_min_prices <- dmv_broadband_now_prices %>%
  mutate(down_minus_25 = ifelse(download >= 25, download - 25, NA)) %>%
  drop_na(down_minus_25) %>%
  mutate(down_minus_100 = ifelse(download <= 100, 100 - download, NA)) %>%
  drop_na(down_minus_100) %>%
  group_by(block_group) %>%
  summarize(price = median(price, na.rm = T), tract = tract) %>%
  group_by(tract) %>%
  summarize(med_price_25 = median(price, na.rm = T)) %>%
  mutate(tract = as.character(tract)) %>%
  full_join(dmv_min_prices, by = "tract")
```

# join with ACS data

```{r}
# join with ACS
dmv_prices <- dmv.tr[substr(dmv.tr$GEOID, 1, 5) %in% c("51013", "51059", "51107", "51510", "51600", "51153", "51683", "51685", "51610", "11001", "24031", "24033", "24017", "24021"),] %>% left_join(dmv_med_and_min_prices, by = c("GEOID" = "tract"))
va_prices <- va.tr %>% left_join(va_med_and_min_prices, by = c("GEOID" = "tract"))
dc_prices <- dc.tr %>% left_join(dc_med_and_min_prices, by = c("GEOID" = "tract"))
md_prices <- md.tr %>% left_join(md_med_and_min_prices, by = c("GEOID" = "tract"))
```

# 

```{r}
ncr.isp.price <- st_drop_geometry(dmv_prices) %>% mutate(year = 2021, region_type = "tract", measure_type = "price", measure_units = "dollars") %>%
  gather(measure, value, c(med_price_25, min_price_25, med_price_100, min_price_100_plus, min_price_100)) %>%
  rename(geoid = GEOID, region_name = NAME) %>%
  select("geoid", "region_type", "region_name", "year", "measure", "value", "measure_type", "measure_units") %>%
  relocate("geoid", "region_type", "region_name", "year", "measure", "value", "measure_type", "measure_units") %>%
  drop_na()

va.isp.price <- st_drop_geometry(va_prices) %>% mutate(year = 2021, region_type = "tract", measure_type = "price", measure_units = "dollars") %>%
  gather(measure, value, c(med_price_25, min_price_25, med_price_100, min_price_100_plus, min_price_100)) %>%
  rename(geoid = GEOID, region_name = NAME) %>%
  select("geoid", "region_type", "region_name", "year", "measure", "value", "measure_type", "measure_units") %>%
  relocate("geoid", "region_type", "region_name", "year", "measure", "value", "measure_type", "measure_units") %>%
  drop_na()

dc.isp.price <- st_drop_geometry(dc_prices) %>% mutate(year = 2021, region_type = "tract", measure_type = "price", measure_units = "dollars") %>%
  gather(measure, value, c(med_price_25, min_price_25, med_price_100, min_price_100_plus, min_price_100)) %>%
  rename(geoid = GEOID, region_name = NAME) %>%
  select("geoid", "region_type", "region_name", "year", "measure", "value", "measure_type", "measure_units") %>%
  relocate("geoid", "region_type", "region_name", "year", "measure", "value", "measure_type", "measure_units") %>%
  drop_na()

md.isp.price <- st_drop_geometry(md_prices) %>% mutate(year = 2021, region_type = "tract", measure_type = "price", measure_units = "dollars") %>%
  gather(measure, value, c(med_price_25, min_price_25, med_price_100, min_price_100_plus, min_price_100)) %>%
  rename(geoid = GEOID, region_name = NAME) %>%
  select("geoid", "region_type", "region_name", "year", "measure", "value", "measure_type", "measure_units") %>%
  relocate("geoid", "region_type", "region_name", "year", "measure", "value", "measure_type", "measure_units") %>%
  drop_na()
```

# send first set of data to db

```{r}
# send to tract level data db
con <- get_db_conn(db_pass = "rsu8zvrsu8zv")
dc_dbWriteTable(con, "dc_digital_communications", "ncr_tr_broadband_now_2021_internet_package_price_update", ncr.isp.price)
dc_dbWriteTable(con, "dc_digital_communications", "va_tr_broadband_now_2021_internet_package_price_update", va.isp.price)
dc_dbWriteTable(con, "dc_digital_communications", "dc_tr_broadband_now_2021_internet_package_price_update", dc.isp.price)
dc_dbWriteTable(con, "dc_digital_communications", "md_tr_broadband_now_2021_internet_package_price_update", md.isp.price)
dbDisconnect(con)
```

# get county, block group, and health district data for DC, MD, and VA using get_data function

```{r}
get_data <- function(va_broadband_now_prices, county = T)
{
  if (county)
  {
    va_100_min_prices <- va_broadband_now_prices %>%
      mutate(down_minus_100 = ifelse(download >= 100, download - 100, NA)) %>%
      drop_na(down_minus_100) %>%
      group_by(block_group) %>%
      slice_max(order_by = -down_minus_100, n = 1) %>%
      slice_max(order_by = -price, n = 1)  %>%
      mutate(county = substr(block_group, 1, 5)) %>%
      group_by(county) %>%
      summarize(min_price_100 = median(price, na.rm = T))
    va_100_min_prices.2 <- va_broadband_now_prices %>%
      mutate(down_minus_100 = ifelse(download >= 100, download - 100, NA)) %>%
      drop_na(down_minus_100) %>%
      group_by(block_group) %>%
      slice_max(order_by = -price, n = 1)  %>%
      mutate(county = substr(block_group, 1, 5)) %>%
      group_by(county) %>%
      summarize(min_price_100_plus = median(price, na.rm = T)) %>%
      full_join(va_100_min_prices, by = "county")
    va_100_med_and_min_prices <- va_broadband_now_prices %>%
      mutate(down_minus_100 = ifelse(download >= 100, download - 100, NA)) %>%
      drop_na(down_minus_100) %>%
      group_by(block_group) %>%
      slice_max(order_by = -down_minus_100, n = 1) %>%
      mutate(county = substr(block_group, 1, 5)) %>%
      group_by(county) %>%
      summarize(med_price_100 = median(price, na.rm = T)) %>%
      full_join(va_100_min_prices.2, by = "county")
    va_min_prices <- va_broadband_now_prices %>% # find minimum price for 25 in each block group, then median price among those in tract
      mutate(down_minus_25 = ifelse(download >= 25, download - 25, NA)) %>%
      drop_na(down_minus_25) %>%
      mutate(down_minus_100 = ifelse(download <= 100, 100 - download, NA)) %>%
      drop_na(down_minus_100) %>%
      group_by(block_group) %>%
      slice_max(order_by = -price, n = 1)  %>%
      mutate(county = substr(block_group, 1, 5)) %>%
      group_by(county) %>%
      summarize(min_price_25 = median(price, na.rm = T)) %>%
      full_join(va_100_med_and_min_prices, by = "county")
    va_med_and_min_prices <- va_broadband_now_prices %>%
      mutate(down_minus_25 = ifelse(download >= 25, download - 25, NA)) %>%
      drop_na(down_minus_25) %>%
      mutate(down_minus_100 = ifelse(download <= 100, 100 - download, NA)) %>%
      drop_na(down_minus_100) %>%
      group_by(block_group) %>%
      summarize(price = median(price, na.rm = T), tract = tract) %>%
      mutate(county = substr(block_group, 1, 5)) %>%
      group_by(county) %>%
      summarize(med_price_25 = median(price, na.rm = T)) %>%
      full_join(va_min_prices, by = "county")
  }
  else
  {
    va_100_min_prices <- va_broadband_now_prices %>%
      mutate(down_minus_100 = ifelse(download >= 100, download - 100, NA),) %>%
      drop_na(down_minus_100) %>%
      group_by(block_group) %>%
      slice_max(order_by = -down_minus_100, n = 1) %>%
      slice_max(order_by = -price, n = 1)  %>%
      group_by(block_group) %>%
      summarize(min_price_100 = median(price, na.rm = T))
    va_100_min_prices.2 <- va_broadband_now_prices %>%
      mutate(down_minus_100 = ifelse(download >= 100, download - 100, NA)) %>%
      drop_na(down_minus_100) %>%
      group_by(block_group) %>%
      slice_max(order_by = -price, n = 1)  %>%
      group_by(block_group) %>%
      summarize(min_price_100_plus = median(price, na.rm = T)) %>%
      full_join(va_100_min_prices, by = "block_group")
    va_100_med_and_min_prices <- va_broadband_now_prices %>%
      mutate(down_minus_100 = ifelse(download >= 100, download - 100, NA)) %>%
      drop_na(down_minus_100) %>%
      group_by(block_group) %>%
      slice_max(order_by = -down_minus_100, n = 1) %>%
      group_by(block_group) %>%
      summarize(med_price_100 = median(price, na.rm = T)) %>%
      full_join(va_100_min_prices.2, by = "block_group")
    va_min_prices <- va_broadband_now_prices %>% # find minimum price for 25 in each block group, then median price among those in tract
      mutate(down_minus_25 = ifelse(download >= 25, download - 25, NA)) %>%
      drop_na(down_minus_25) %>%
      mutate(down_minus_100 = ifelse(download <= 100, 100 - download, NA)) %>%
      drop_na(down_minus_100) %>%
      group_by(block_group) %>%
      slice_max(order_by = -price, n = 1)  %>%
      group_by(block_group) %>%
      summarize(min_price_25 = median(price, na.rm = T)) %>%
      full_join(va_100_med_and_min_prices, by = "block_group")
    va_med_and_min_prices <- va_broadband_now_prices %>%
      mutate(down_minus_25 = ifelse(download >= 25, download - 25, NA)) %>%
      drop_na(down_minus_25) %>%
      mutate(down_minus_100 = ifelse(download <= 100, 100 - download, NA)) %>%
      drop_na(down_minus_100) %>%
      group_by(block_group) %>%
      summarize(price = median(price, na.rm = T), tract = tract) %>%
      group_by(block_group) %>%
      summarize(med_price_25 = median(price, na.rm = T)) %>%
      full_join(va_min_prices, by = "block_group")
  }
  return(va_med_and_min_prices)
}

health_district <- read.csv("/project/biocomplexity/sdad/projects_data/vdh/va_county_to_hd.csv")
health_district$county_id <- as.character(health_district$county_id)

va_bg_broadband_now_prices2 <- left_join(health_district, va_broadband_now_prices, by = c("county_id" = "county"))
va_100_min_prices <- va_bg_broadband_now_prices2 %>%
  mutate(down_minus_100 = ifelse(download >= 100, download - 100, NA),) %>%
  drop_na(down_minus_100) %>%
  group_by(block_group) %>%
  slice_max(order_by = -down_minus_100, n = 1) %>%
  slice_max(order_by = -price, n = 1)  %>%
  group_by(health_district) %>%
  summarize(min_price_100 = median(price, na.rm = T))
va_100_min_prices.2 <- va_bg_broadband_now_prices2 %>%
  mutate(down_minus_100 = ifelse(download >= 100, download - 100, NA)) %>%
  drop_na(down_minus_100) %>%
  group_by(block_group) %>%
  slice_max(order_by = -price, n = 1)  %>%
  group_by(health_district) %>%
  summarize(min_price_100_plus = median(price, na.rm = T)) %>%
  full_join(va_100_min_prices, by = "health_district")
va_100_med_and_min_prices <- va_bg_broadband_now_prices2 %>%
  mutate(down_minus_100 = ifelse(download >= 100, download - 100, NA)) %>%
  drop_na(down_minus_100) %>%
  group_by(block_group) %>%
  slice_max(order_by = -down_minus_100, n = 1) %>%
  group_by(health_district) %>%
  summarize(med_price_100 = median(price, na.rm = T)) %>%
  full_join(va_100_min_prices.2, by = "health_district")
va_min_prices <- va_bg_broadband_now_prices2 %>% # find minimum price for 25 in each block group, then median price among those in tract
  mutate(down_minus_25 = ifelse(download >= 25, download - 25, NA)) %>%
  drop_na(down_minus_25) %>%
  mutate(down_minus_100 = ifelse(download <= 100, 100 - download, NA)) %>%
  drop_na(down_minus_100) %>%
  group_by(block_group) %>%
  slice_max(order_by = -price, n = 1)  %>%
  group_by(health_district) %>%
  summarize(min_price_25 = median(price, na.rm = T)) %>%
  full_join(va_100_med_and_min_prices, by = "health_district")
va_med_and_min_prices.hd <- va_bg_broadband_now_prices2 %>%
  mutate(down_minus_25 = ifelse(download >= 25, download - 25, NA)) %>%
  drop_na(down_minus_25) %>%
  mutate(down_minus_100 = ifelse(download <= 100, 100 - download, NA)) %>%
  drop_na(down_minus_100) %>%
  group_by(block_group) %>%
  summarize(price = median(price, na.rm = T), tract = tract) %>%
  mutate(county_id = substr(block_group, 1, 5)) %>%
  right_join(health_district, by = c("county_id")) %>%
  group_by(health_district) %>%
  summarize(med_price_25 = median(price, na.rm = T)) %>%
  full_join(va_min_prices, by = "health_district")

# county
va_med_and_min_prices.ct <- get_data(va_broadband_now_prices)
dc_med_and_min_prices.ct <- get_data(dc_broadband_now_prices)
md_med_and_min_prices.ct <- get_data(md_broadband_now_prices)
dmv_med_and_min_prices.ct <- get_data(dmv_broadband_now_prices)

# block group
va_med_and_min_prices.bg <- get_data(va_broadband_now_prices, F)
dc_med_and_min_prices.bg <- get_data(dc_broadband_now_prices, F)
md_med_and_min_prices.bg <- get_data(md_broadband_now_prices, F)
dmv_med_and_min_prices.bg <- get_data(dmv_broadband_now_prices, F)
```

# get block and county level data initial formatting

```{r}
dmv.bg <- get_acs(geography = "block group",
                 year = 2019,
                 variables = c(median_household_income = "B19013_001"),
                 state = c("VA", "DC", "MD"),
                 survey = "acs5",
                 output = "wide",
                 geometry = TRUE)

dmv_prices.ct <- dmv.ct[dmv.ct$GEOID  %in% c("51013", "51059", "51107", "51510", "51600", "51153", "51683", "51685", "51610", "11001", "24031", "24033", "24017", "24021"),] %>% left_join(dmv_med_and_min_prices.ct, by = c("GEOID" = "county"))
dmv_prices.bg <- dmv.bg[substr(dmv.bg$GEOID, 1, 5) %in% c("51013", "51059", "51107", "51510", "51600", "51153", "51683", "51685", "51610", "11001", "24031", "24033", "24017", "24021"),] %>% left_join(dmv_med_and_min_prices.bg, by = c("GEOID" = "block_group"))

va_prices.ct <- va.ct %>% left_join(va_med_and_min_prices.ct, by = c("GEOID" = "county"))
va_prices.bg <- va.bg %>% left_join(va_med_and_min_prices.bg, by = c("GEOID" = "block_group"))

dc_prices.ct <- dc.ct %>% left_join(dc_med_and_min_prices.ct, by = c("GEOID" = "county"))
dc_prices.bg <- dc.bg %>% left_join(dc_med_and_min_prices.bg, by = c("GEOID" = "block_group"))

md_prices.ct <- md.ct %>% left_join(md_med_and_min_prices.ct, by = c("GEOID" = "county"))
md_prices.bg <- md.bg %>% left_join(md_med_and_min_prices.bg, by = c("GEOID" = "block_group"))
```

# finalize formatting

```{r}
ncr.isp.price.ct <- st_drop_geometry(dmv_prices.ct) %>%
  mutate(year = 2021, region_type = "county", measure_type = "price", measure_units = "dollars") %>%
  gather(measure, value, c(med_price_25, min_price_25, med_price_100, min_price_100_plus, min_price_100)) %>%
  rename(geoid = GEOID, region_name = NAME) %>%
  select("geoid", "region_type", "region_name", "year", "measure", "value", "measure_type", "measure_units") %>%
  relocate("geoid", "region_type", "region_name", "year", "measure", "value", "measure_type", "measure_units") %>%
  drop_na()
ncr.isp.price.bg <- st_drop_geometry(dmv_prices.bg) %>%
  mutate(year = 2021, region_type = "block group", measure_type = "price", measure_units = "dollars") %>%
  gather(measure, value, c(med_price_25, min_price_25, med_price_100, min_price_100_plus, min_price_100)) %>%
  rename(geoid = GEOID, region_name = NAME) %>%
  select("geoid", "region_type", "region_name", "year", "measure", "value", "measure_type", "measure_units") %>%
  relocate("geoid", "region_type", "region_name", "year", "measure", "value", "measure_type", "measure_units") %>%
  drop_na()

va.isp.price.ct <- st_drop_geometry(va_prices.ct) %>%
  mutate(year = 2021, region_type = "county", measure_type = "price", measure_units = "dollars") %>%
  gather(measure, value, c(med_price_25, min_price_25, med_price_100, min_price_100_plus, min_price_100)) %>%
  rename(geoid = GEOID, region_name = NAME) %>%
  select("geoid", "region_type", "region_name", "year", "measure", "value", "measure_type", "measure_units") %>%
  relocate("geoid", "region_type", "region_name", "year", "measure", "value", "measure_type", "measure_units") %>%
  drop_na()
va.isp.price.bg <- st_drop_geometry(va_prices.bg) %>%
  mutate(year = 2021, region_type = "block group", measure_type = "price", measure_units = "dollars") %>%
  gather(measure, value, c(med_price_25, min_price_25, med_price_100, min_price_100_plus, min_price_100)) %>%
  rename(geoid = GEOID, region_name = NAME) %>%
  select("geoid", "region_type", "region_name", "year", "measure", "value", "measure_type", "measure_units") %>%
  relocate("geoid", "region_type", "region_name", "year", "measure", "value", "measure_type", "measure_units") %>%
  drop_na()

dc.isp.price.ct <- st_drop_geometry(dc_prices.ct) %>%
  mutate(year = 2021, region_type = "county", measure_type = "price", measure_units = "dollars") %>%
  gather(measure, value, c(med_price_25, min_price_25, med_price_100, min_price_100_plus, min_price_100)) %>%
  rename(geoid = GEOID, region_name = NAME) %>%
  select("geoid", "region_type", "region_name", "year", "measure", "value", "measure_type", "measure_units") %>%
  relocate("geoid", "region_type", "region_name", "year", "measure", "value", "measure_type", "measure_units") %>%
  drop_na()
dc.isp.price.bg <- st_drop_geometry(dc_prices.bg) %>%
  mutate(year = 2021, region_type = "block group", measure_type = "price", measure_units = "dollars") %>%
  gather(measure, value, c(med_price_25, min_price_25, med_price_100, min_price_100_plus, min_price_100)) %>%
  rename(geoid = GEOID, region_name = NAME) %>%
  select("geoid", "region_type", "region_name", "year", "measure", "value", "measure_type", "measure_units") %>%
  relocate("geoid", "region_type", "region_name", "year", "measure", "value", "measure_type", "measure_units") %>%
  drop_na()

md.isp.price.ct <- st_drop_geometry(md_prices.ct) %>%
  mutate(year = 2021, region_type = "county", measure_type = "price", measure_units = "dollars") %>%
  gather(measure, value, c(med_price_25, min_price_25, med_price_100, min_price_100_plus, min_price_100)) %>%
  rename(geoid = GEOID, region_name = NAME) %>%
  select("geoid", "region_type", "region_name", "year", "measure", "value", "measure_type", "measure_units") %>%
  relocate("geoid", "region_type", "region_name", "year", "measure", "value", "measure_type", "measure_units") %>%
  drop_na()
md.isp.price.bg <- st_drop_geometry(md_prices.bg) %>%
  mutate(year = 2021, region_type = "block group", measure_type = "price", measure_units = "dollars") %>%
  gather(measure, value, c(med_price_25, min_price_25, med_price_100, min_price_100_plus, min_price_100)) %>%
  rename(geoid = GEOID, region_name = NAME) %>%
  select("geoid", "region_type", "region_name", "year", "measure", "value", "measure_type", "measure_units") %>%
  relocate("geoid", "region_type", "region_name", "year", "measure", "value", "measure_type", "measure_units") %>%
  drop_na()
```

# send almost all data to db (except for tract level - which I did before adding the rest)

```{r}
# send to db
con <- get_db_conn(db_pass = "rsu8zvrsu8zv")
dc_dbWriteTable(con, "dc_digital_communications", "ncr_ct_broadband_now_2021_internet_package_price", ncr.isp.price.ct)
dc_dbWriteTable(con, "dc_digital_communications", "va_ct_broadband_now_2021_internet_package_price", va.isp.price.ct)
dc_dbWriteTable(con, "dc_digital_communications", "dc_ct_broadband_now_2021_internet_package_price", dc.isp.price.ct)
dc_dbWriteTable(con, "dc_digital_communications", "md_ct_broadband_now_2021_internet_package_price", md.isp.price.ct)

# probably want to change the naming here...
dc_dbWriteTable(con, "dc_digital_communications", "ncr_bg_broadband_now_2021_internet_package_price_update", ncr.isp.price.bg)
dc_dbWriteTable(con, "dc_digital_communications", "va_bg_broadband_now_2021_internet_package_price_update", va.isp.price.bg)
dc_dbWriteTable(con, "dc_digital_communications", "dc_bg_broadband_now_2021_internet_package_price_update", dc.isp.price.bg)
dc_dbWriteTable(con, "dc_digital_communications", "md_bg_broadband_now_2021_internet_package_price_update", md.isp.price.bg)
dbDisconnect(con)

# read in health district name
con <- get_db_conn(db_pass = "rsu8zvrsu8zv")
health_district_geoids <- st_read(con, query = "SELECT * FROM dc_geographies.va_hd_vdh_2021_health_district_geo_names")
DBI::dbDisconnect(con)

va.isp.price.hd <- merge(va_med_and_min_prices.hd, health_district_geoids, by.x = "health_district", by.y = "region_name") %>%
  mutate(year = 2021, region_type = "health district", measure_type = "price", measure_units = "dollars") %>%
  gather(measure, value, c(med_price_25, min_price_25, med_price_100, min_price_100_plus, min_price_100)) %>%
  rename(region_name = health_district) %>%
  select("geoid", "region_type", "region_name", "year", "measure", "value", "measure_type", "measure_units") %>%
  relocate("geoid", "region_type", "region_name", "year", "measure", "value", "measure_type", "measure_units") %>%
  drop_na()

con <- get_db_conn(db_pass = "rsu8zvrsu8zv")
dc_dbWriteTable(con, "dc_digital_communications", "va_hd_broadband_now_2021_internet_package_price", va.isp.price.hd)
dbDisconnect(con)
```
